### 2019-12-19 manuscript.tex

\documentclass[crop]{CSLB}

\usepackage{MySettings}

\input{captions}

\begin{document}

\bibliographystyle{mn2e}

\supertitle{Research Paper}

\title[Probabilities of SETI contacts]{Probability of causal
contact between interstellar civilizations through MonteCarlo simulations}

\author[Lares \& Funes]{Lares M$^{1, 2}$, Funes J$^{1, 3}$}

\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}

\address{
   %
   \add{1}{CONICET, Argentina}

   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
   Astron\'omico de C\'ordoba, Argentina}

   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
   %
}

\begin{abstract}
%
The abundance of intelligent civilizations in the galaxy is a
longstanding question, which is often conceptualized as the problem
of the lack of received communication or the Fermi paradox.
%
Most efforts on the estimation of a number of intelligent
civilizations are centered on the Drake equation, although 
its factors are affected by large uncertainties, and it lacks a
temporal nature.
%
A different approach uses detailed numerical simulations of the galaxy
and recipes for the rates of the formation of stars and planets, and
even for the origin of life.
%
Here we present numerical simulations of stochastic processes of
emergence of civilizations with communication capabilities, based
on minimal asumptions, and analyze the causal connections among them.
%
The analysis of the rate of causal contacts as a function of the mean
number of civilizations, the mean lifetime span distribution and the
maximum distance a civilization can send signals is presented, and
used as a framework to discuss the spatial and temporal structure
of a populated galaxy within several different scenarios.
%
Our results indicate that, given the large distances involved, causal
contacts between civilizations are very rare.
%
Additionally, the odds to make a contact in a few years of monitoring, 
assuming a perfect detection rate, are low for most models, with the
exception of models which propose a galaxy densely populated with
long lived civilizations.
%
The probability of causal contacts increases with the increase of the 
mean lifetime of civilizations much more signifficantly than with
the increase in the mean number of active civilizations for a given
time window.
%
\end{abstract}

\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
causal contact between civilizations through MonteCarlo simulations,
International Journal of Astrobiology}

\keywords{SETI, Computer simulations, Statistics}

\JELclassification{XX; XX}
\MSCcodes{XX; XX}

\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
Discrete Event, GHZ: Galactic Habitable Zone}

\maketitle

\section{Motivations}\label{S_motivations}
%{{{

A common approach to discussing the estimations of the number of
extraterrestrial intelligent civilizations (ETI) in the Galaxy has
been the use of the Drake equation \citep{gleiser_drake_2010,
prantzos_joint_2013, haqq-misra_drake_2017}.
%
The abundance of intelligent civilizations in the galaxy is a
longstanding question, which is often conceptualized as the problem of
the lack of received communication or the Fermi paradox
\citep{barlow_galactic_2012, Sotos_biotechnology_2019,
forgan_galactic_2016}.
%
There are many proposals aimed at solving this paradox, which make use
of statistical \citep{solomonides_probabilistic_2016,
vanhouten_isthere_2017, horvat_calculating_2007,
maccone_statistical_2015} or stochastic approaches
\citep{forgan_numerical_2009, bloetscher_using_2019,
glade_stochastic_2011, forgan_numerical_2010}, analytical
interpretations of the Drake equation \citep{prantzos_joint_2013,
smith_broadcasting_2009} or more speculative proposals
\citep{barlow_galactic_2013, lampton_information_2013,
conway_three_2018, forgan_galactic_2017}.
%
The uncertainties in the factors of the Drake equation make it less
prone to a formal application in order to define searching strategies
or to compute the actual number of ETIs, but is a key tool to organize
the discussion around the problem of the population of ETIs in the
Galaxy \citep{hinkel_interdisciplinary_2019}, which is noticeable
dominated by the fact that despite decades of effort not a single ETI,
if any, has been detected so far.
%
Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, or to propose a probabilistic approach,
or to consider the temporal structure which is missing in the
equation.
%

\Fpagebreak

Temporal aspects of the distribution of communicating civilizations
and their contacts have also been explored by several authors
\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011},
%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
%The spatiotemporal aspects of SETI
%
as well as efforts on considering the stochastic nature of the Drake equation
\citep{glade_stochastic_2011}.
%
The simulation approach has also benn considered
\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017},
although with a similar problem that consists on the large number of
parameters which are either unknown or largely uncertain.
%
%New numerical determination of habitability in the Galaxy: the SETI connection
%
%Since the factors in the Drake equation are uncertain,
%
In this work, we propose to
avoid the frequentist approach of the Drake equation and to explore a
parameter space, where instead of computing a final number, we provide
a statistical distribution that gives conditional probabilities.
%
This is an exploratory analysis that aims at providing a numerical 
tool to discuss not the several theoretical problems summarized by the
Drake equation factors, 
but the different scenarios on the basis of statistical heuristics.
%
The approach proposed here shold be considered as a compromise between
the uncertainties of the frequentist approach and the detailed recipes
required on the simulation approach.
%
We then provide a numerical framework to explore via simulations a
parameter space of unknown observables in order to discuss different
scenarios and their consecuences in terms of the probability con
making contacts.
%
The only fact that can be stated with certainty is that for the number
of years SETI projects have been working we have not received any
signal, at least
within the technical posibilities and the 
conditions stablished by SETI \citep{tarter_search_2001}.
%
This arguably astonishing result has lead to the Fermi paradox, which
states an apparent contradiction between the high estimation of the
abundance of life in the Galaxy and the lack of their detection
\citep{vanhouten_isthere_2017}.
%
The lack of detection has also ispired to a number of alternative proposals
for new SETI strategies \citep{forgan_exoplanet_2017,
balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.

The discrete events method for simulating a stochastic process is an
approximation that allows to study the behaviour of complex
systems, by considering a sequence of well defined discrete events.
%
The simulation os carried out by following all the variables that
describe the system, that constitute the state of the system.
%
The evolution of the process, then, is described as a set of changes
in the state of the system.
%
In this context, an event produces a specific change in the state,
that can be triggered by random variables that encode the stochastic
nature of the physical phenomenon.
%
The process involves following the changes on the state of the system,
definig the initial and final states, defining a method that allows to
keep track of time progress, and maintaining a list of relevant
events.

In a recent work, \citep{balbi_impact_2018} use a statistical model to analyze
the occurrence of causal contacts between civilizations in the Galaxy.
%
The author highlights the effect of evolutionary processes when
attempting to estimate the number of communicating civilizations that
might be in causal contact with an observer located on the Earth.
%
\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
structure in the Drake equation.

%While the Drake equation remains a very useful
%guide in identifying and analyzing the various factors in-
%volved in the problem and their relative probabilities, our
%approach could provide a better framework with which to
%perform a statistical analysis when taking into account evo-
%lutionary processes. In particular, we have shown that the
%Drake equation can misestimate the number of detectable
%communicating species if spatial and temporal dependencies
%are important. This should suggest some caution in the un-
%critical application of such an equation. In this paper, we offer
%an exemplificative illustration of the effect of different
%probability distributions for the typical timescales involved
%in the problem. Ideally, it would be desirable to arrive at a
%mathematical description of such distributions based on a
%modeling of the underlying astrophysical and biological
%processes. This points out a possible future direction of in-
%vestigation in which an integrated evolutionary (and coevo-
%lutionary) approach to the problem would be a central focus
%of interest (for more on this perspective, see, e.g., Chaisson

\citet{bloetscher_using_2019} uses a probabilistic approach, though
still strongly motivated by the Drake equation, to obtain a
propbabilistic measure for the number of civilizations in the Galaxy.
To that end, perform MonteCarlo Markov Chains over each factor of the
Drake equation, and combine the means to obtain a probabilistic
result. It is worth mentioning that the author proposes a lof-normal
target distribution to compute the posterior probabilities. This
approach, which assigns probability distributions to the Drake
factors. This study concludes that there is a very low probability for
Earth to make contact with another extraterrestrail civilization.

\citet{smith_broadcasting_2009} uses a analytical model to compute the
probabilities of contact between two randomly located civilizations,
and the waiting time for the first contact, assuming a fixed maximum
broadcasting distance.

\citet{balbi_impact_2018} investigate the chance of communicating
civilizations making causal contact within a volume surrounding the
location of the Earth, using a statistical model to compute the
occurrence of causal contacts. The author stresses the fact that the
causal contact requirement involves the distance between
civilizations, their lifespan and their times of appeareance.  This is
important since the time it takes the light to travel across the
Galaxy is much lesser than the lifetime of the Galaxy.

\citet{balbi_impact_2018} fix the total number of civilizations and
explore the parameter space that comprise three parameters, namely,
the distance to the Earth, the time of appeareance and the lifespan of
the communicating civilization. Each of these three variables are
drawn from a random distribution. The distribution of the distances
results from a uniform distribution of civilization within the plane
of the Galaxy. For the distribution of the characteristic time of
appeareance, the author explores and exponential and a trucated
Gaussian distributions. For the distribution of the lifespans, the
author choses an exponential distribution. The radius of influence is
set to 1000 ly.

\citet{balbi_impact_2018} finds that the fraction of communicating
civilizations vary with the choice of the statistical distribution for
the time of appeareance. For all analyzed distributions, the fraction
corresponding to a mean lifetime of 10$^9$ years and for a maximu
radius for the detection of the signal of 1000 ys, is roughly 0.1.

\citet{vukotic_astrobiological_2012} propose a simulation approach
based on probabilistic cellular automata (PCA) modeling. In this
framework, a complex system is modeled by a lattice of cells which
evolve at discrete time steps, acording to transition rules that take
into account for each cell the states of its neighbour cells.

The authors implement a PCA model to a network of cells which
represent life complexity on a 2-dimensional annular ring resembling
the GHZ. The authors set the GHZ between a minimum radius of 6 kpc and
a maximum radius of 10 kpc. Within this framework, the authors also
make several MonteCarlo simulations and analyze ensemble-averaged
results.

This work aims at analyzing the evolution of life, although it does
not account for the network of causal contacts among technological
civilizations.

%- - - -

In this work we address the problem of the temporal and spatial
structure of the distribution of communicating civilizations, by
exploring the hypothesis space over a minimal set of parameters.
%
In Sec.~\ref{S_methods} we introduce the methods and discuss the
candidate distributions for the statistical aspects of the times
involved in the communication process.
%
Then, we present our results in Sec~\ref{S_results}, with special
emphasis on the statistical distributions of the duration of causal
contacts in one or both directions, the possible differences on the
position of a civilization on the Galaxy and the distribution of time
intervals for the waiting of the first contact, always as a function
of the three simulation parameters.
%
Finally, in Sec.~\ref{S_discussion} we discuss our results and future
research directions.

%}}}

% HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}

\input{methods}

\input{results}

\section{Discusion}\label{S_discussion}
%{{{

We have implemented a suite of simulations following a stochastic
approach, to explore the hypothesis space of a simple model that
accounts for the causal connections between communicating
civilizations on an simplified galaxy.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a Galaxy where an intelligent civilization is
very rare, to a Galaxy populated with plenty of civilizations in
causal contact.
%
We explored the consecuences of different models using mensurable
quantities, that arise as by-products of the simulations.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space.
%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios for
the spread of intelligent life on the galaxy.

Although the simulations use a number of speculative assumptions, we
argue that the degree of knowledge about the origin and persistence of
life in the Galaxy does not worth the implementation of more detailed
or sophisticated models.
%
Instead, we take advantadge of the simplicity of the model to explore
the hypothesis space, i.e., the three--dimensional parameter space of
the parametric model, in order to gain insight on the consecuences of
different scenarios for the search of intelligent life.
%
Our analysis is not centered in obtaining the odds for the Earth
making contact with another extraterrestrial civilization, but in
obtaining a statistical, parameter dependent description of the
communication system consisting of a group of broadcasting
civilizations which arise sparsely both in time and space in the
Galaxy.

Under the hypotheses of our experiments, we conclude that a causal
contact is extremely unlikely unless the Galaxy is heavily populated
by intelligent civilizations.
%
Even so, we must consider that in order to stablish a contact between
any two entities, a minimum degree of compatibility must be
accomplished without any previous agreement, making the posibility of
a contact with a message that could be deciphered highly rare.

Taking into account the caveat mentioned previously about the
trade--off between the simplicity and complexity of the experiment,
future work can be performed following this framework in order to
explore possible implications of the results for more detailed
configurations.
%
For example, the communication method (isotropic, colimated,
serendipitous) can affect the observables.
%
Since we analyze the causal contacts, effective contacts depend on the
efficiency of both the emitter and the receiver.
%
A correction by coverage ratio in the detection and by a targetting
ratio in the emission could be easily implemented in the simulation,
although the effect would be to reduce the probability of contact by
an efficiency ratio.
%
Thus, such implementation is not really necessary, since the values of
probabilities could modified by a constant correction factor equal to
the combined emission/reception efficiency
\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
forgan_collimated_2014}.
%
If we consider a statistical distribution for this factor, then the
results could change, and this could be studied with the DE approach.
%
Also the nature of the message carrier, electromagnetic or another.
%
It is also possible to explore the effects of alignments or the use of
stars as sources or amplifiers.
%
%energetic markers Spiritual markers
%
As a future work, we will explore several improvements to the
simulation, and analyze the impact on the final results of these
changes.
%
They include a spatial distribution that actually resembles the shape
of the Galaxy, different distribution functions for the mean lifetime
of MPLs and different efficiencies for the sphere of the causal region
of each MPL, resembling different searching strategies
\citep{hippke_interstellar_2017}.

Possible hypothesis improvements are:

\begin{itemize}
   \item D$_{max}$ is different for different MPLs.  For example a power
      law where powerfull emmisions are rare and weak emssions are
      common.
   \item The probability of the rising of a new MPL vary within the GHZ.
   \item Correction by coverage ratio in the detection and by
      targetting ratio in the emission.
   \item Spiritual factor.  Explore the role of the S factor: message
      contents, influence on the lifespan of a MPL that receives a
      message, mean lifespan of a MPL that emits a message.
\end{itemize}

%}}}

\ack[Acknowledgement]{
%
This work was partially supported by the Consejo Nacional de
Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
Argentina.
%
This research has made use of NASA's Astrophysics Data System.
%
Plots and simulations were made with software developed by the authors
using R and python languajes. Plots were postprocessed with inkscape.
%
}

\bibliography{biblio_seti}

\end{document}