### 2019-12-19 results.tex

\section{Results: exploring the parameter space}\label{S_results}
%{{{

We implemented the simulation of a regular grid of models varying over
the hypothesis space, which covers 15000 different models.
%
For each model, we simulated 50 realizations with different random
seeds, in order to estimate the variance.
%
The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
4000-200000 and 10000-500000, respectively, with a partition of 50
values for each parameter.
%
For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
20000, 40000 and 80000 lyr.
%
This setup makes a total of 750000 simulations, each covering a time
range of one million years.
%
In Table \ref{T_simu_hypotheses} we show the range of values of these parameters explored 
in our numerical experiments.

As a product of the simulations, several quantities are obtained.
%
Some quantities are directly derived from the discrete events, namely,
the ID of emitting MPL, the ID of emitting (receiving) MPL, the
position in the galaxy, the time of Awakening (Contact) and the time
of Doomsday  (Blackout) for each MPL.
%
This structure is always for the A-D cycle of each emitting MPL, and
repeats for each contact (C-B).
%
We can also derive the additional quantities.
%
The age of a MPL is the time elapsed from A to D.
%
%
%given time, ceti\_e is the MPL emitting signals (emiter), ceti\_r is
%the MPL receiving signals (receiver), ceti\_c is the ceti that
%listen at least another ceti (citizen) and ceti\_h is the MPL ceti
%that is lestened by at least another MPL.
%
The time span of a MPL listening another or being listened by another
can also be easily derived from the results of a realization.
%
This way we can also compute the distribution in the galaxy of cetis
that reach contact and the waiting time until the first contact or the
waiting time until the next contact.
%
Regarding the properties of the population of MPLs and its evolution,
we analyze the fraction of awaken time a ceti is listening at least
another ceti, the age of contacted MPLs at first contact, the
distribution of time to wait until next contact, the fraction of cetis
where the first contact is given at the awakening, the distribution of
the number of contacts for each MPL, the distribution of the maximum
number of contacts for each MPL, the distribution of the number of
contacts as a function of MPL age, the number of contacts as a
function of time in the galaxy, the rate of MPLs that succeed in
contact, and the distribution of distances between contacted MPLs.
%
Usefull derived quantity is the duration of two-way communication
channels the or the fraction of contacts that admit a response.
% 
We also analyze the relations between the distance to ceti vs. time of
double communication, the distance to ceti vs. age of contacted ceti,
the age of a ceti and the maximum number of contacted cetis before D,
or the lifespan of a ceti vs. max number of contacts.
%
All these quantities can be analyzed as a function of the simulation
parameters.
%
We chose eight models that are on fairly oposed regions of the
explored parameter space, and cover short/long lifetimes, short/long
waiting times for the next MPL to appear in the Galaxy and
short/large range reach of the signal, including all possible
combinations.
%
The details of each model are summarized in the
Table~\ref{T_selected_models}.

In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
distributions of the number contacts for MPLs in four different
samples (M5, M6, M7, M8) with long range signal reach
(D$_{max}$=10000). 
%
The lifetime is more determinant than the awakening rate in increasing
the number of contacts.
%
As expected, of the four samples, the sample M7, which corresponds to
a dense awakening in the timeline and a long lifetime, is the model
that maximizes the number of contacts, reaching a maximum of order 100
contacts for a single MPL.
%
This is considerable larger than any model with a shorter lifetime,
which produce a number of contacts of at most the order of ten
contacts per MPL.
%
A couple of comments are worth to mention about this plot.
%
First, it has a logarithmic scale on the x-axis, and second it is the
cumulative, not differential, empirical distribution.

In what follows, we focus on the analysis of the duration of contacts
(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
(Sec.~\ref{SS_waiting}) and the
variations of contact posibilities as a function of the position of
the MPL in the Galaxy (Sec.~\ref{SS_location}).

%}}}

\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
%{{{

The duration of each contact for a single receiving MPL is given by the
time difference between a C event and a B event, which are direct
outputs for each simulation experiment.

%}}}

\subsection{Waiting time for a first contact}\label{SS_waiting}
%{{{

In this section we analyze the distribution of the waiting times for
a first contact.
%
%Such distribution can be considered to compute the probability of
%reaching a first contact within the first XX years of a SETI project,
%assuming the methods are correct and the hypotheses of the experiment.
Such distribution can be considered to compute the probability for a
random MPL to carry out a SETI project and spend a given time until
the first contact is made.

In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the MPLs
that have made at least one contact as a funtion of the elapsed time
since the awakening, given that they make at least one contact on
their lifetime, for different models.
%
From a frequentist approach, the normalized cumulative counts are
related to the estimation of the probability of observing a given time
with no success.
%             
Both panels show the histograms of the waiting times for the first
contact, for all the MPLs that make at least one contact along their
lifetime.
%
We show separately the histograms for several values of $\tau_a$ and a
fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
and a fixed value of $(\tau_a, D_{max})$.
%
For different values of $\tau_a$, the shape of the distribution is
similar, and the main difference given by the different number of
contacts on the different models.
%
Conversely, for different values of $\tau_a$, the shape of the
distribution is similar, and the main difference given by the
different number of contacts on the different models. 
%
We show the 2D maps of the rate of MPLs that never make contact and
the rate of MPLs that make contact at the moment of the awakening,
inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
as a function of $\tau_s$ and $\tau_s$.
%
In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
waiting times for the first contact, for several models.
%
Left panel shows the histograms for several values of $\tau_a$, and
$\tau_s$ in the range 17000-24000 yr.
%
Right panel shows the histograms for several values of $\tau_s$, and
$\tau_a$ in the range 68000-100000 yr.
%
The shape of the histograms does not change when $\tau_a$ is modified,
but significantly changes when $\tau_s$ is modified.

%}}}

\subsection{Location within the Galaxy}\label{SS_location}
%{{{

%On the Figure~\ref{F_radial_frac} we show the fraction of MPLs that
%make zero, one, two or three contacts, as a function of the
%galactocentric distance, for model M1.  VERIFICAR!  M1?
%
For this model, the majority of MPLs do not make any contact during
all the time period in which they are active.
%
As expected, the fractions of MPLs making at least one contact
diminishes as the number of contacts rises.
%
The shape of this distribution is a result of the geometry of the GHZ.
%
In general, a closer position to the center of the Galaxy slightly
favours the posibility of a contact.
%
This would indicate that the position of the Earth is suitable for the
odds of making a contact.

%}}}

\setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{ccccc}
\hline
Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
\hline
M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
\hline
\end{tabular}
\caption{Selected models to analize the behavior of simulation outputs
   and their dependence on simulation outputs and their dependence on
   simulation parameters.}
\label{T_selected_models}
\end{table*}