### 2019-12-22 manuscript.tex
con cambios propuestos por Jose

\documentclass[crop]{CSLB}

\usepackage{MySettings}

\input{captions}

\begin{document}

\supertitle{Research Paper}

\title[Probabilities of SETI contacts]{Probability of causal
contact between interstellar civilizations through MonteCarlo simulations}

\author[Lares, Funes \& Gramajo]{Lares M$^{1, 2}$, Funes J.~G.$^{1, 3}$ \& Gramajo
L$^{1, 2}$}

\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}

\address{
   \add{1}{CONICET, Argentina}

   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
           Astron\'omico de C\'ordoba, Argentina}

   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
}

\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
causal contact between civilizations through MonteCarlo simulations,
International Journal of Astrobiology}

\keywords{SETI, Computer simulations, Statistics}

\JELclassification{XX; XX}
\MSCcodes{XX; XX}

\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
Discrete Event, GHZ: Galactic Habitable Zone}

\begin{abstract}
%
The abundance of intelligent civilizations in the galaxy is a
longstanding question, which is often conceptualized as the problem
of the lack of received communication or the Fermi paradox.
%
Most efforts on the estimation of a number of intelligent
civilizations are centered on the Drake equation, although 
its factors are affected by large uncertainties, and it lacks a
temporal nature.
%
A alternative approach uses detailed numerical simulations of the galaxy
and recipes for the rates of the formation of stars and planets, and
even for the origin of life.
%
Here we present numerical simulations of stochastic processes of
emergence of civilizations with communication capabilities, based
on minimal asumptions, and analyze the causal connections among them.
%
The analysis of the rate of causal contacts as a function of the mean
number of civilizations, the mean lifetime span distribution and the
maximum distance a civilization can send signals is presented, and
used as a framework to discuss the spatial and temporal structure
of a populated galaxy within several different scenarios.
%
Our results indicate that, given the large distances involved, causal
contacts between civilizations are very rare.
%
Additionally, the odds to make a contact in a few years of monitoring, 
assuming a perfect detection rate, are low for most models, with the
exception of models which propose a galaxy densely populated with
long lived civilizations.
%
The probability of causal contacts increases with the 
mean lifetime of civilizations much more signifficantly than with
the mean number of active civilizations for a given
time window.
%
\end{abstract}

\maketitle

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\input{motivations}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\input{methods}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\input{results}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Discusion}\label{S_discussion}
%{{{

We have implemented a suite of simulations following a stochastic
approach, to explore the hypothesis space of a simple model that
accounts for the causal connections between communicating
civilizations on an simplified galaxy.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a Galaxy where an intelligent civilization is
very rare, to a Galaxy populated with plenty of civilizations in
causal contact.
%
We explored the consecuences of different models using mensurable
quantities, that arise as by-products of the simulations.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space.
%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios for
the spread of intelligent life on the galaxy.

Although the simulations use a number of speculative assumptions, we
argue that the degree of knowledge about the origin and persistence of
life in the Galaxy does not worth the implementation of more detailed
or sophisticated models.
%
Instead, we take advantadge of the simplicity of the model to explore
the hypothesis space, i.e., the three--dimensional parameter space of
the parametric model, in order to gain insight on the consecuences of
different scenarios for the search of intelligent life.
%
Our analysis is not centered in obtaining the odds for the Earth
making contact with another extraterrestrial civilization, but in
obtaining a statistical, parameter dependent description of the
communication system consisting of a group of broadcasting
civilizations which arise sparsely both in time and space in the
Galaxy.

Under the hypotheses of our experiments, we conclude that a causal
contact is extremely unlikely unless the Galaxy is heavily populated
by intelligent civilizations.
%
Even so, we must consider that in order to stablish a contact between
any two entities, a minimum degree of compatibility must be
accomplished without any previous agreement, making the posibility of
a contact with a message that could be deciphered highly rare.

Taking into account the caveat mentioned previously about the
trade--off between the simplicity and complexity of the experiment,
future work can be performed following this framework in order to
explore possible implications of the results for more detailed
configurations.
%
For example, the communication method (isotropic, colimated,
serendipitous) can affect the observables.
%
Since we analyze the causal contacts, effective contacts depend on the
efficiency of both the emitter and the receiver.
%
A correction by coverage ratio in the detection and by a targetting
ratio in the emission could be easily implemented in the simulation,
although the effect would be to reduce the probability of contact by
an efficiency ratio.
%
Thus, such implementation is not really necessary, since the values of
probabilities could modified by a constant correction factor equal to
the combined emission/reception efficiency
\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
forgan_collimated_2014}.
%
If we consider a statistical distribution for this factor, then the
results could change, and this could be studied with the DE approach.
%
Also the nature of the message carrier, electromagnetic or another.
%
It is also possible to explore the effects of alignments or the use of
stars as sources or amplifiers.
%
%energetic markers Spiritual markers
%
As a future work, we will explore several improvements to the
simulation, and analyze the impact on the final results of these
changes.
%
They include a spatial distribution that actually resembles the shape
of the Galaxy, different distribution functions for the mean lifetime
of MPLs and different efficiencies for the sphere of the causal region
of each MPL, resembling different searching strategies
\citep{hippke_interstellar_2017}.

Possible hypothesis improvements are:

\begin{itemize}
   \item D$_{max}$ is different for different MPLs.  For example a power
      law where powerfull emmisions are rare and weak emssions are
      common.
   \item The probability of the rising of a new MPL vary within the GHZ.
   \item Correction by coverage ratio in the detection and by
      targetting ratio in the emission.
   \item Explore the role of the message
      contents, influence on the lifespan of a MPL that receives a
      message, mean lifespan of a MPL that emits a message.
\end{itemize}

%}}}

\ack[Acknowledgement]{
%
This work was partially supported by the Consejo Nacional de
Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
Argentina.
%
This research has made use of NASA's Astrophysics Data System.
%
Plots and simulations were made with software developed by the authors
using R and python languajes. Plots were postprocessed with inkscape.
%
}

\bibliographystyle{mn2e}
\bibliography{biblio_seti}

\end{document}