### 2020-01-13 manuscript.tex

\documentclass[crop]{CSLB}

\usepackage{MySettings}

\input{captions}

\begin{document}

\supertitle{Research Paper}

\title[Probabilities of SETI contacts]{Probability of causal
contact between interstellar civilizations through MonteCarlo simulations}

\author[Lares, Funes \& Gramajo]{Lares M.$^{1, 2}$, Funes J.~G.$^{1,
3}$ \& Gramajo L.$^{1, 2}$}

\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}

\address{
   \add{1}{CONICET, Argentina}

   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
           Astron\'omico de C\'ordoba, Argentina}

   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
}

\keywords{SETI, Computer simulations, Statistics}

%\JELclassification{XX; XX}
%\MSCcodes{XX; XX}

\Abbreviations{SETI: Search for Extraterrestrial Intelligence,\\
CCN: causally connected node,\\
SFC: surface of first contact,\\
SLC: surface of last contact,\\
DE: Discrete Event,\\
GHZ: Galactic Habitable Zone}

\begin{abstract}
%
The abundance of intelligent civilizations in the Galaxy is a
   longstanding question, which is often conceptualized as the problem
   of the lack of received communication or the Fermi paradox.
%
Most efforts on the estimation of a number of intelligent
   civilizations are focused on the Drake equation, although its
   factors are affected by large uncertainties, and it lacks a
   temporal nature.
%
An alternative approach uses detailed numerical simulations of the
   galaxy and recipes for the rates of the formation of stars and
   planets, and even for the origin of life.
%
These two opposite approaches rely on a large number of highly
   uncertain parameters.
%
Here we present numerical simulations of stochastic processes for the
   emergence of nodes with communication capabilities and analyze the
   causal connections among them.
%
This is based on minimal assumptions and three free parameters, with a
   focus on the statistical properties of empirical population models
   instead of an interpretable model with variables to be determined
   by observation.
%
The analysis of the rate of causal contacts as a function of the mean
   number of civilizations, the mean lifespan distribution and the
   maximum distance a civilization can send signals is presented, and
   used as a framework to discuss the spatial and temporal structure
   of a populated galaxy within several different scenarios.
%
Our results indicate that, given the large distances involved, causal
   contacts between civilizations are very rare.
%
Additionally, the odds to make a contact in a few years of monitoring,
   assuming a perfect detection rate, are low for most models, with
   the exception of models which propose a galaxy densely populated
   with long lived civilizations.
%
The probability of causal contacts increases with the mean lifetime of
   civilizations much more significantly than with the mean number of
   active civilizations for a given time window.
%
Finally, we show that maximum probability of making a contact occurs
when a civilization acquires the required communication technology.
\end{abstract}

\maketitle

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\input{Sec_motivations}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\input{Sec_methods}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\input{Sec_results}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Discussion}\label{S_discussion}
%{{{

We have implemented a suite of simulations following a stochastic
approach, to explore the hypothesis space of a simple model that
accounts for the causal connections between communicating
civilizations on a simplified galaxy.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a galaxy model where an intelligent civilization is
very rare, to a galaxy model populated with plenty of civilizations in
causal contact.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space,
so we explored the outcomes of different models 
that arise as by-products of the simulations.
%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios for
the appearance of intelligent life on the Galaxy.

Although the simulations use a number of speculative assumptions, we
argue that the degree of knowledge about the origin and persistence of
life in the Galaxy does not worth the implementation of more detailed
or sophisticated models.
%
Instead, we take advantage of the simplicity of the model to explore
the hypothesis space, i.e., the three--dimensional parameter space of
the parametric model, in order to gain insight on the consequences of
different scenarios for the search of intelligent life.
%
Our analysis is not centered in obtaining the odds for the Earth
to make contact with another intelligent civilization.
%
Instead, we focus on obtaining a statistical, parameter dependent
description of the possible properties of the communication networks
that comprise sets of nodes with broadcasting and reception
capabilities.
%
This causally connected nodes are sparsely distributed in both space
and time, making it difficult the analytical treatment and justifying
the simulation approach.

Under the hypotheses of our experiments, we conclude that a causal
contact is extremely unlikely unless the galaxy is heavily populated
by intelligent civilizations with large average lifetimes.
%
Roughly, in order to have at least one contact in the entire lifetime of the
\ceti{}, there should appear a mean of at least one \ceti{} every $\sim$20000 yr, with
a mean active period of at least $\sim$30000 yr.
%
There is a balance between the density of \cetis{} and the mean
lifetime since, as expected, a lower density can be compensated by a
longer active time period.
%
However, a large number of nodes does not easily compensate their
short lives 
to reach the same probability of causal contact than in the case of a
less populated galaxy but with very ancient civilizations.
%
In all cases, for a short period of time (for instance, the time SETI
programs have been active on Earth), the maximum probability of making
a contact occurs at the moment of the awakening.
%
This suggests the possibility that an alternative SETI strategy could be
the search for alternative message carriers, for the case in which the
search has not been performed on the adequate channels.
%
Then, if a contact is produced for the first time, the origin of the
signal is more likely to be very old.
%
Also, the chances of entering the causal connected zone of a \ceti{}
does not grow linearly, but favours the first period of 100000 years.

In the approach of this work, we used computer simulations to address
the problem of the probabilities of causal contacts between location
in the galaxy the possibility of sending and receiving messages.
%
Instead of making a number of assumptions, we have explored the
hypothesis space, reducing the problem to only three parameters and a
few simple hypothesis to perform a complete model for the population and
communication network in the galaxy.
%
This allows to consider the Fermi paradox from a new perspective, and
to propose an alternative treatment for the number of intelligent
emitter/receivers with emphasis on the temporal dimension which limits
the probabilities of contacts due to the short time interval for
rise and fall of civilizations compared to the age and extension of
our Galaxy.

%}}}

\ack[Acknowledgement]{
%
This work was partially supported by the Consejo Nacional de
Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
Argentina.
%
This research has made use of NASA's Astrophysics Data System.
%
Plots and simulations were made with software developed by the authors
using R and python languajes. Plots were postprocessed with inkscape.
%
}

\ack[Disclosure statement]{
%
No competing financial interests exist.
%
}

\bibliographystyle{mn2e}
\setlength{\bibsep}{0.0pt}

\bibliography{biblio_seti}

\end{document}