### 2020-05-05 submit_v01/manuscript.tex

\documentclass[crop]{CSLB}

%The authors can define any packages after the \documentclass{CSL} command.
\usepackage{amsmath} %for dealing with mathematics,
\usepackage{amsthm} %for dealing with theorem environments,
\usepackage{cite} %%%for dealing with citations
\usepackage{hyperref} %for linking the cross references
\usepackage{graphics} %for dealing with figures.
\usepackage{algorithmic} %for describing algorithms
\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
\usepackage{url} %It provides better support for handling and breaking URLs.

%\usepackage{natbib}
\usepackage{boites}

%\usepackage{graphicx}
%\usepackage{float}
%\usepackage{psfig}
%\usepackage{amsfonts,amssymb}
%\usepackage{color}
%\usepackage{textcomp}
%\usepackage[T1]{fontenc}
%\usepackage{aecompl}
%\usepackage{times}
%\usepackage{txfonts}
%\usepackage[english]{babel}
%\usepackage[utf8x]{inputenc}
%\usepackage{amsmath}
%\usepackage{caption}
%\usepackage{siunitx}

\jname{Astrobiology}%

\newtheorem{theorem}{Theorem}
\newtheorem{condition}{Condition}

\newcommand{\ceti}{CCN}
\newcommand{\cetis}{CCNs}

\newcommand{\awakening}{A event }
\newcommand{\contact}{C event }
\newcommand{\blackout}{B event }
\newcommand{\doomsday}{D event }

\newcommand{\aawakening}{Awakening event }
\newcommand{\ccontact}{Contact event }
\newcommand{\bblackout}{Blackout event }
\newcommand{\ddoomsday}{Doomsday event }

\newcommand{\awakenings}{A events }
\newcommand{\contacts}{C events }
\newcommand{\blackouts}{B events }
\newcommand{\doomsdays}{D events }

\newcommand{\ff}{{\color{red}\rule{10pt}{10pt}}}

%\newcommand{\ffn}[1]{ \colorbox{red}{FIG {#1}} }    
\newcommand{\ffn}[1]{}

%\newcommand{\ttn}[1]{ \colorbox{green}{TABLE {#1}} }    
\newcommand{\ttn}[1]{}

\input{captions}

\begin{document}

\supertitle{Research Paper}

\title[Probabilities of SETI contacts]{Probability of causal
contact between interstellar civilizations through MonteCarlo simulations}

\author[Lares, Funes \& Gramajo]{Lares M.$^{1, 2}$, Funes J.~G.$^{1,
3}$ \& Gramajo L.$^{1, 2}$}

\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}

\address{
   \add{1}{CONICET, Argentina}

   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
           Astron\'omico de C\'ordoba, Argentina}

   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
}

\keywords{SETI, Computer simulations, Statistics}

%\JELclassification{XX; XX}
%\MSCcodes{XX; XX}

\Abbreviations{SETI: Search for Extraterrestrial Intelligence,\\
CCN: causally connected node,\\
SFC: surface of first contact,\\
SLC: surface of last contact,\\
DE: Discrete Event,\\
GHZ: Galactic Habitable Zone}

\begin{abstract}
%
The abundance of intelligent civilizations in the Galaxy is a longstanding
question, often conceptualized as the problem of the lack of received
communication. Most efforts on the estimation of a number of civilizations are
focused on the Drake equation, although its factors are affected by large
uncertainties and it lacks a temporal nature.  Alternative approaches use
detailed numerical simulations of the galaxy and recipes for the stars and
planets formation rates and for the origin of life, and rely on highly
uncertain parameters. We present a statistical model for the abundance and
duration of civilizations, implemented through Monte Carlo simulations.  We
explore the hypothesis space of the model by a suite of simulations for the
emergence of communicating nodes, and analyze their
causal connections.  This is based on minimal assumptions and three
free parameters, with focus on the statistical properties of empirical
models instead of an interpretable model with variables to be determined by
observation. The analysis of the rate of causal contacts as a function of the
mean number of civilizations, the mean lifespan distribution and the maximum
distance a civilization can send signals is presented, and used as a framework
to discuss the spatial and temporal structure of a populated galaxy within
several scenarios. We find that, given the large distances involved,
causal contacts between civilizations are very rare. Additionally, the odds to
make a contact in a few years of monitoring are low for most models, with the
exception of those of a galaxy densely populated with long lived
civilizations. The probability of causal contacts increases with the
lifetime of civilizations much more significantly than with the number of
active civilizations for a given time window. Finally, we show that the
maximum probability of making a contact occurs when a civilization acquires
the required communication technology.
%
\end{abstract}

\maketitle

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Motivations}\label{S_motivations}
%{{{

The Drake equation \citep{drake_intelligent_1962} offers a very
helpful educated guess, a rational set of lenses --the factors in the
equation-- through which to look at future contacts with a
technologically advanced civilization in the Milky Way.
%
The equation quantifies the number of civilizations from whom we might
receive an electromagnetic signal, using a number of factors that have
been widely discussed in the literature and whose estimates are
updated constantly.
%
A comprehensive review and analysis of each term of the equation is
presented in \citet{vakoch_drake_2015}.
%
Some optimistic estimates from the Drake equation contrast with the so
called Fermi paradox, a statement about the contradiction between the
apparent abundance of life in the Galaxy and its lack of evidence
\citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
barlow_galactic_2012, forgan_galactic_2016, vanhouten_isthere_2017,
Sotos_biotechnology_2019, carroll_nellemback_fermi_2019}.
%
There are many proposals aimed at solving this paradox, which make use
of statistical \citep{solomonides_probabilistic_2016,
vanhouten_isthere_2017, horvat_calculating_2007,
maccone_statistical_2015} or stochastic approaches
\citep{forgan_numerical_2009, bloetscher_using_2019,
glade_stochastic_2011, forgan_numerical_2010}.
%
Regarding the Drake equation, analytical interpretations
\citep{prantzos_joint_2013,
smith_broadcasting_2009} or reformulations
\citep[][and references therein]{burchell_whither_2006} have also been
proposed.
%
The lack of contacts could be explained by the difficulty of life to
appear due to astrophysical explanations
\citep{annis_astrophysical_1999}.
%
In general, also more speculative proposals
\citep{barlow_galactic_2013, lampton_information_2013,
conway_three_2018, forgan_galactic_2017} have been discussed.
% 
\Fpagebreak
%
We stress the fact that the only observation that can be stated with
certainty is that for the number of years projects for the Search of
Extraterrestrial Intelligences (SETI) have been working, we have not
received any signal using the technical possibilities and the
conditions established by SETI projects \citep{tarter_search_2001}.
%
The lack of detection has also inspired to a number of alternative
proposals for new SETI strategies \citep{forgan_exoplanet_2017,
balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
%
The Drake equation is a key tool to organize the discussion around the
problem of the abundance of civilizations in the Galaxy
\citep{hinkel_interdisciplinary_2019}.
%
However, the uncertainties in the factors of this equation make it
less prone to a formal application in order to define searching
strategies or to compute the actual number of extraterrestrial
intelligences.
%
The optimistic estimations from the Drake equation also contrast with
the SETI initiatives, noticeable overshadowed by the fact that despite
decades of effort, not a single extraterrestrial intelligence, if any,
has been detected.
%
\citet{prantzos_joint_2013} proposes a unified framework for a joint
analysis of the Drake equation and the Fermi paradox, concluding that
for sufficiently long-lived civilizations, colonization of the Galaxy
is the only reasonable option to gain knowledge about other life
forms.
%
\citet{haqq-misra_drake_2017} discuss the dependence of the Drake
equation parameters on the spectral type of the host star and the time
since the Galaxy formed and examine trajectories for the emergence of
communicative civilizations over the history of the Galaxy.
%
Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, to propose a probabilistic approach, or
to consider the temporal structure which is missing in the equation.
%
Temporal aspects of the distribution of communicating civilizations
and their contacts have also been explored by several authors
\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
well as efforts on considering the stochastic nature of the Drake
equation \citep{glade_stochastic_2011}.
%
The simulation approach has also been considered
\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017,
ramirez_new_2017}, although with a similar problem that consists on
the large number of parameters which are either unknown or largely
uncertain.
%
\citet{grimaldi_signal_2017} consider a statistical model for the
probability of the Earth making contact with other intelligent
civilizations, taking into account the finite lifetime of signal
emitters, and based on the fractional volume occupied by all signals
reaching our planet.

In a recent work, \citep{balbi_impact_2018} use a statistical model to
analyze the occurrence of causal contacts between civilizations in the
Galaxy.
%
The author highlights the effect of evolutionary processes when
attempting to estimate the number of communicating civilizations that
might be in causal contact with an observer located on the Earth.
%
\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
structure in the Drake equation.
%
\citet{bloetscher_using_2019} uses a probabilistic approach, though
still strongly motivated by the Drake equation, to obtain a
probabilistic measure for the number of civilizations in the Galaxy.
%
To that end, perform MonteCarlo Markov Chains over each factor of the
Drake equation, and combine the means to obtain a probabilistic
result.
%
It is worth mentioning that the author proposes a log-normal target
distribution to compute the posterior probabilities.
%
%This approach, which assigns probability distributions to the Drake
%factors.
%
This study concludes that there is a very low probability for Earth to
make contact with another extraterrestrial civilization.    
%
\citet{smith_broadcasting_2009} uses an analytical model to compute
the probabilities of contact between two randomly located
civilizations, and the waiting time for the first contact, assuming a
fixed maximum broadcasting distance.
% 
Using a different approach, \citet{balbi_impact_2018} investigate the
chance of communicating civilizations making causal contact within a
volume surrounding the location of the Earth, using a statistical
model to compute the occurrence of causal contacts.
%
The author stresses the fact that the causal contact requirement
involves the distance between civilizations, their lifespan and their
times of appearance.  This is important since the time it takes the
light to travel across the Galaxy is much lesser than its lifetime.
%
\citet{balbi_impact_2018} fix the total number of civilizations and
explore the parameter space that comprise three parameters, namely,
the distance to the Earth, the time of appearance and the lifespan of
the communicating civilization.
%
Each of these three variables are drawn from a random distribution.
The distribution of the distances results from a uniform distribution
of civilization within the plane of the Galaxy.
%
For the distribution of the characteristic time of appearance, the
author explores and exponential and a truncated Gaussian
distributions. For the distribution of the lifespans, the author
chooses an exponential distribution.
%
%The radius of influence is set to 1000 lyr.
%
\citet{balbi_impact_2018} finds that the fraction of communicating
civilizations vary with the choice of the statistical distribution for
the time of appearance. For all analyzed distributions, the fraction
corresponding to a mean lifetime of 10$^9$ years and for a maximum
radius for the detection of the signal of 1000 yr, is roughly 0.1.
%
In another probabilistic approach,
\citet{vukotic_astrobiological_2012} propose a probabilistic cellular
automata (PCA) modeling. In this framework, a complex system is
modeled by a lattice of cells which evolve at discrete time steps,
according to transition rules that take into account for each cell the
states of its neighbor cells.
%
The authors implement a PCA model to a network of cells which
represent life complexity on a 2-dimensional annular ring resembling
the Galactic Habitable Zone (GHZ). 
%
The authors set the GHZ between a minimum radius of 6 kpc and a
maximum radius of 10 kpc. Within this framework, the authors also make
several MonteCarlo simulations and analyze ensemble-averaged results.
%
This work aims at analyzing the evolution of life, although it does
not account for the network of causal contacts among technological
civilizations.

In this work we address the problem of the temporal and spatial
structure of the distribution of communicating civilizations, by
exploring the hypothesis space over a minimal set of parameters.
%
We propose to avoid the frequentist approach of the
Drake equation and to explore a parameter space, where instead of
computing a final number, we provide a statistical distribution that
gives conditional probabilities.
%
This is an exploratory analysis that aims at providing a numerical
tool to discuss not the several theoretical problems summarized by the
Drake equation factors, but the different scenarios on the basis of
statistical heuristics.
%
The approach proposed here should be considered as a compromise
between the uncertainties of the frequentist approach and the detailed
recipes required on the simulation approach.
%
We then provide a numerical framework to explore via simulations a
parameter space of unknown observables in order to discuss different
scenarios and their consequences in terms of the probability of
making contacts.
%
The method we use for simulating a stochastic process is an
approximation that allows to study the behavior of complex systems,
by considering a sequence of well defined discrete events.
%
The simulation is carried out by following all the variables that
describe and constitute the state of the system, and the evolution of
the process is described as a set of changes in that state.
%
In this context, an event produces a specific change in the state,
that can be triggered by random variables that encode the stochastic
nature of the physical phenomenon.
%
For example, when a new contact is produced between two entities in
the simulated galaxy, the numbers of active communication lines and of
active communicated nodes change.
%
The process involves following the changes on the state of the
system, defining the initial and final states, defining a method that
allows to keep track of the time progress in steps, and maintaining a
list of relevant events.
%
In Sec.~\ref{S_methods} we introduce the methods and discuss the
candidate distributions for the statistical aspects of the times
involved in the communication process.
%
Then, we present our results in Sec~\ref{S_results}, with special
emphasis on the statistical distributions of the duration of causal
contacts in one or both directions, the possible differences on the
position of a civilization on the Galaxy and the distribution of time
intervals for the waiting of the first contact, always as a function
of the three simulation parameters.
%
Finally, in Sec.~\ref{S_discussion} we discuss our results and future
research directions.

%}}}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Methods and working hypotheses}\label{S_methods}
%{{{

Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
An advantage of simulations compared to theoretical approaches, is
that the former usually require less assumptions and simplifications,
and can be applied to systems where a theoretical model can not be
found.
%
In particular, many complex stochastic processes that can be described
by the evolution of the state of a system, can be efficiently modeled
with the discrete--event (hereafter, DE) simulation approach.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a timeline and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
ross_simulation_2012}.
%
Each event yields modifications to the values of the variables that define
the process, producing the corresponding change in the state of the
simulated system.
%
This method is well suited for the particular case of the diffusion of
intelligent signals in the galaxy, and allows to explore several
models easily.
%
We simulate the statistical properties of a set of points in space and
time that have a causal connection at light speed, hereafter dubbed
''Causal Contact Nodes'' (hereafter, \ceti{}).
%
We chose this generic name in order to stress the fact that in this
analysis no broadcasting or lookout activities are required, and
therefore not actual messages are considered, only the causal contact.
%
For the special case of a fully efficient node that emits and receives
isotropically, a causal contact node can be considered as a
broadcasting station that has the ability to detect signals through an
active lookout program.
%
Also, it is worth mentioning that this is a general approach, and not
necessarily a \ceti{} is the host of an intelligent civilization.
%
For instance, it can be associated with a planet where life has developed, became
intelligent, reached the skills required to find the right
communication channel, sustain a search and establish a contact.
%
Alternative message processing entities could be considered, for
example interstellar beacons where life has ceased to exist but
continue with its emission, or communication stations established by
probes or left intentionally by intelligent beings \citep[see, e.g.,
][]{peters_outer_2018, barlow_galactic_2013}.
%
In principle, these strategies could affect our results since it would
be easier to configure a cluster of \cetis{} that spread in time.
%
However, we do not consider these highly speculative alternatives at
this point.
%
For the purpose of this analysis, only the communication capability is
relevant, since we study the causal contacts between the locations.
%
The system is defined by a number of actors that represent \cetis{}
and appear at different instants in time, generating events that
produce meaningful changes in the variables that describe the system,
i.e., in the arrangement of \cetis{} and their network of causal
contacts.
%
For example, the appearance of a new \ceti{} in a region which is
filled with a signal emitted by another node, will increase the
number of active \cetis{} and the number of pairs of \cetis{} in
causal contact.
%
Assuming some simple hypotheses, the discrete events method can be
performed taking into account a small number of variables, which allow
to analyze the variation of the results in the model parameter space.

In what follows, we describe the experimental setup chosen to estimate
the probabilities of causal contacts and several derived quantities in
terms of three independent parameters, namely, the mean time span
between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
their mean lifetime $\left<\tau_s\right>$, and the maximum distance a
signal can be detected by another \ceti{} ($D_{max}$).
%
Intuitively, the shorter the $\tau_a$ parameter and the larger the
$\tau_s$ and $D_{max}$ parameters are, the greater the probability of
the existence of causal contacts between pairs of \cetis{} would be.
%
We also need to propose theoretical distribution functions for both
the distribution of lifespans ($\tau_s$) and the distribution of the
number of \cetis{} per unit time \citep{maccone_evolution_2014,
Sotos_biotechnology_2019}, related to the time span between the
appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$
is shorter, it produces a greater density of \cetis{}).
%
The shapes of these distributions are set to a fixed law, as discussed
in Sec.~\ref{SS_PDF_shape}.

%}}}

\begin{figure}[!t]
   \centering
   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
   \usecaption{F_scheme} \label{F_scheme}
\end{figure}

%{{{

\ffn{1} We illustrate in the Fig.~\ref{F_scheme} the schematic
representation of the region causally connected to a given \ceti{}.
%
To this end, we use space--time diagrams, where time is represented on
the horizontal axis, and space is represented in the vertical axis.
%
For the sake of simplicity, we show in the plot the light travel
distance, i.e., the distance traversed by any signal spreading from
the emitter at light speed (for example, any electromagnetic signal).
%
In this scheme, a light pulse would follow a trajectory represented by
a line at 45 degrees from the axes, given the units of the time and
space axes are years and light years, respectively.
%
Panel (a) of the Figure represents the region $R$ (shaded polygon) in
space and time which is causally connected to the emitter (left
vertex), following an event (hereafter A type event) in which the node
acquires the communication capability, or becomes ''active''.
%
The sphere of first contact (SFC) is a sphere centered on the emitter,
represented in the Figure by the left angle of the region $R$.
%
This sphere grows until its radius reaches the D$_{max}$ distance, in
which it would be no longer detectable due to the decrease in the
energy per unit area, which falls under the assumed detectability
threshold.
%
Similarly, the surface of last contact (SLC) is another sphere that
grows from an event in which the nodes ceases to possess the
communication capability (D type event), and carries the last signal
produced by the emitter.
%
The region which is causally connected to the emitter is then limited
by these two spheres, and has the shape of a sphere (before the D
event) or of a spherical shell (after the D event).
%
The temporal intervals for the communication between two \cetis{} are
represented in the panel (b) of the Figure.
%
In this scheme, the entire lifetime of a \ceti{} can then be
represented by a polygon, limited in the spatial axis by the maximum
distance of the signal $D_{max}$ and in the time axis by the time span
between the starting time (A event, concave vertex) and the ending
time (D event, convex vertex).
%
This representation allows to easily visualize the important events
that result from the existence and communications of two \cetis{}.
%
The points in time where the \cetis{} acquire their communicating
capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
Figure).
%
Similarly, the points in time where the \cetis{} loss their
communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
in the Figure).
%
The points in space--time where the first contact is produced for each
one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
and $C_2$ in the Figure.
%
Finally, the two points where the contact is lost for each one the
\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
$B_2$ in the Figure.
%
The receiver node E$_1$ can listen signals from the emitter node
E$_2$, from the first contact event (t=C$_{12}$) up to the last
contact event (t=B$_{12}$).
%
Similarly, the receiver node E$_2$ can listen signals from the emitter
node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
%
The time intervals for the open communication channel are determined
by the ''C'' and ''B'' type events.
%
It is important to notice that there is a time delay between the
contact events of the two \cetis{} involved in this analysis, and also
a time delay between the two blackout events, so that the time range
when a bidirectional contact is possible occurs between the maximum
time of the contact points and the minimum time between the blackout
points.

%}}}

\begin{figure*}
   \centering
   \includegraphics[width=\textwidth]{F_messages.pdf}
   \usecaption{F_messages} \label{F_messages}
\end{figure*}

%{{{

The temporal structure that emerges naturally from the experimental
setup implies that,
as a fundamental property of the simulations,
a causal connection can be produced without the need of the two
involved locations being concurrent or active at the same time.
%
This property of the system arises as a
consequence of the large spatial and temporal scales, where a message
is transmitted at the (relatively small) light speed.
%
Although a \ceti{} could be active for enough time so as to transmit a
message at large distances, the limited power of the message and the
dilution that runs as the squared distance from the source, imposes a
detectability limit.
%
As a consequence of this limitation and of their finite lifetime,
considered as the period of time between the acquisition and loss of
communicating capacity, each \ceti{} will fill a spherical shell
region of the galaxy, limited by two concentric spherical surfaces.
%
The leading front, or surface of first contact (SFC) grows from the
central \ceti{} until it reaches the maximum distance $D_{max}$.
%
Thus, the volume reached causally by a \ceti{} is initially a growing
sphere.
%
Following the end of the civilization, there is still a region which
is filled with the emitted signals.
%
This approach has been also considered other statistical models
\citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
%
The trailing front, or surface of last contact (SLC) also grows from
the central \ceti{}, with a delay with respect to the SFC equivalent
to the lifetime of the node, and produces a spherical shell region.
%
Any other \ceti{} within this region will be aware of the originating
node, even if it has disappeared before the time of contact.
%
This region will grow if the surface of first contact has not yet
reached the maximum distance $D_{max}$, and will shrink otherwise
until the surface of last contact reaches $D_{max}$, producing as a
result the loss of all signal from the central \ceti{}.
%
In our approach, we consider a model galaxy where the width of the
disk is negligible with respect to the radius of the disk.
%
In the 2D simulation only the intersection of the communicating
spherical shells with the plane of the galaxy is relevant, and produce
the corresponding circles or rings for the filled spheres or annular
regions of the spherical shells, respectively.
%
The initially growing communicating sphere, is shown over space--time
diagrams, where time is represented on the horizontal direction, and
space is represented in the vertical direction.
%
In the Fig.~\ref{F_scheme} the two
emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
%
The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
%
Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
%
The type and length of causal contact in both directions depend on the
distance and time lag between the awakening events, the maximum
distance that a signal can reach and the time period in which each
emitter is active.
%
\ffn{2}
%
In the Fig.~\ref{F_messages} we show a schematic representation of the
possible cases in which a \ceti{} E can be in causal contact with a
\ceti{} R.
%
The duration of the causal contact in one direction depends on several
factors, mainly the lifetimes, the time period between the A events
and the distance between the two nodes involved in the causal
contact.
%
As it can be seen, the direction of the causal contact does not imply
any ordering for the A events of the involved nodes.
%
In the implementation of the DE simulation, the six cases shown in the
Figure determine the succession of events.
%
Each time an event is produced, the list of upcoming events is updated
according to these cases.

In out experimental configuration, there is no a temporal window for
the model, i.e., the simulation starts when the process is already
stable, and ends before any galactic effect could modify or affect of
fixed variables.
%
Also, we assume that the probability for the rise of a \ceti{} is
homogeneous over the GHZ.
%
Although the Galaxy has a well known spiral structure, the assumed
sparsity of nodes (intelligent civilizations) does make this
assumption fairly reasonable.
%
Otherwise, if the distribution of nodes is not sparse, it could be the
case that the spiral arms would host most of the \cetis{}, and
contacts are frequent between closely located nodes.
%
In such a case we could expect that the interpretation of our results
could be considered as pessimist for the establishment of a contact
with our planet.
%
We also limit the possibilities of life or other types of
civilizations to the hypothesis usually stated for the definition of
the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001,
lineweaver_galactic_2004, gonzalez_habitable_2005,
morrison_extending_2015, haqq-misra_evolution_2019,
rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and
consider that habitability remains constant \citep[see, however,
][]{gonzalez_habitable_2005, dayal_habitability_2016, gobat_evolution_2016}.
%
For example, \citet{vukotic_astrobiological_2012} propose
probabilistic cellular automata model to explain the astrobiological
history of the Milky Way.
%
This means that we set aside possible civilizations that could survive
in severe conditions or unstable systems, which would prevent the
appearance of life as we know it.

While the aforementioned assumptions are basic conditions for most of
the stochastic processes observed in nature, we need to make stronger
assumptions related to the nature of the message.
%
The most simple assumption about the message itself, is that it
travels at light speed.
%
With this, we are considering messages sent through, for example,
electromagnetic radiation or even gravitational radiation or neutrino
communication \citep{hippke_interstellar_2017, wright_how_2018}, but we set
apart messages sent with mechanical means or physical objects
\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology
that violates the known laws of physics.
% 
For the communication of \cetis{} through messages sent isotropically,
we assume that the capacity to emit signals and to receive signals
occur at the same time.
%
This means that the ability to find another civilization develops at
the same time that the ability and intention to emit a signal in all
directions, intended to be detected by an unknown civilization.
%
Although there are several reasons to think that this could not be the
case, at large time scales it can be considered that both abilities
occur roughly at coincident epochs.
%
Another essential assumption is that all \cetis{} use the same signal
power, so that there is a maximum distance out to which it can be
detected.
%
It is worth mentioning that we are considering in our experimental
setup a system composed by several emitters and receivers across the
galaxy, under the hypothesis proposed before.
%
In such a system we compute probabilities of a \ceti{} making contact
with another node, and the results must be interpreted as for a
randomly chosen \ceti{}, given certain conditions of homogeneity on an
ensemble of equivalent galaxies.
%
Then, we stress the fact that we are not centered on the probabilities
for the Earth itself, but on a generalized \ceti{}.
%
Regarding the extent of the signals, we know that the distance from
which a signal from Earth could be detected using the current
technology is of a few hundreds light years, given that the signal was
sent to a specific direction.
%
%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
%[from https://en.wikipedia.org/wiki/Fermi_paradox]
%
It is straightforward to propose and implement a distribution of
maximum distances, although this would increase the model complexity
at the cost of a larger uncertainty or of another dimension on the
hypothesis space.
%
%This proposal is based on a pronoid scenario, and we do not consider
%in this work the occurrence of paranoid or partially paranoid
%civilizations.
%%
%If that would be the case, the results obtained here can be taken as
%upper limits.

In our simulations we also assume the simplest situations for the
growth of the sphere of first contact.
%
For example, we discard the possibility of stellar colonization
\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
maccone_mathematical_2011} and assume that the communication aim is
performed to all directions with the same power and the same
probability, i.e., strongly assume isotropic communication in all
cases.
%
More detailed simulations could be produced considering different
efficiency of communication or detection methods.
%
However, in this case we also would complicate the experimental setup,
make the results less evident.
%
The results are independent of the nature of life (organic or
artificial).
%
The lifetime of a civilization can be caused by auto destruction or by
external factors.
%}}}

 \setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{cllll}
\hline
   \multicolumn{2}{l}{independent variables (hypothesis space)}
   &min. value&max. value&Nbins\\
\hline
   $\tau_{a}$ & Mean temporal separation between consecutive
   awakenings & 4000 yr & 200\,000 yr & 50\\ 
   $\tau_{s}$ & Mean lifetime of a \ceti{}
   & 10000 yr & 500\,000 yr& 50\\ 
   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{(500,
   1000, 10000, 20000, 40000, 80000) lyr} & 6  \\
\hline
   \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
\hline
   & statistical properties of all \cetis{} &equally distributed&&\\
   & Point process for the distribution in time & homogeneous &&\\
   f$_s$ & The scan of the sky & fully efficient&&1\\
   f$_p$ & panspermia or colonization &absent&&0\\
   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
   R$_{GHZ}^{min}$   & Inner radius of the GHZ     & & &20\,000 lyr\\
   R$_{GHZ}^{max}$   & Outer radius of the GHZ     &&&60\,000 lyr\\
   t$_{max}$ & Time span of the simulation  &&&1.e6 yr\\
    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
\hline
   \multicolumn{2}{l}{discrete events} &
   \multicolumn{3}{l}{affected variables}\\
\hline
   A event & Awakening: a \ceti{} starts its communication
   capabilities &Number of active \cetis{}\\
   B event & Blackout: the end of the communication channel stops
   &Number causal channels\\
   C event & Contact: a new causal contact is produced &Number of causal channels\\
   D event & Doomsday: a \ceti{} ends its communication capabilities&Number of active \cetis{}\\
\hline

\hline
\end{tabular}
\caption{Definition of independent variables and adopted values for 
   fixed parameters 
   that are part of the simulation.  Variable parameters define the
   spatial and temporal structure
   of the process and the maximum reach of the messages.}
\label{T_simu_hypotheses}
\end{table*}

\subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
%{{{

The temporal structure of the process is defined by two distribution
parameters that represent the mean time interval that a \ceti{}
can emit and receive signals, and the mean time interval
between the emergence of consecutive \cetis{}.
%
The spatial structure of the simulation is given by the size and shape
of the Galactic Habitable Zone and the maximum distance a signal can
travel to be detected (D$_{max}$).
%
The parameters for the temporal distributions also determine the
spatial properties, since the density of active \cetis{} in the galaxy
depend on these two parameters.
%
For example, a small $\tau_a$ and a large $\tau_s$ will produce a
galaxy densely populated with \cetis{}.
%
Also, some hypotheses regarding the shapes of the distributions of the
temporal parameters must be made in order to complete the
simulation.
%
\citet{forgan_spatiotemporal_2011} argues that the times at which
different civilizations become intelligent follow a Gaussian
distribution, and then the distribution of inter-arrival times is an
inverse exponential.
%
We choose to assume that the distribution of the times of A events is a
stationary Poisson process, and then the distribution of the times
between the appearance of new consecutive \cetis{} is exponential.
%
Regarding the duration of a \ceti{}, we propose that the distribution
of the duration of \cetis{} is a stationary exponential distribution.
%
\citet{balbi_impact_2018} stress the fact that it would be desirable
to arrive at a theoretical statistical distribution of the lifespan of
civilizations, preferentially on the basis of the underlying
astrophysical and biological processes.
%
\citet{maccone_lognormals_2014} argues that this distribution should
be a log normal.

The power law and exponential statistical distributions are among the
most common patterns found in natural phenomena.
%
For example, the distribution of the frequency of words in many
languages is known to follow a Zipf's law (which is a power law).
%
This distributions arise any time a phenomena is characterized by
commonly occurring small events and rarely produced large events
\citep[e.g. ][]{adamic_zipf_2000}.
%
Zipf law also describes population ranks of cities in various
countries, corporation sizes, income rankings, ranks of number of
people watching the same TV channel.
%
The magnitudes of earthquakes, hurricanes, volcanic eruptions and
floods; the sizes of meteorites or the losses caused by business
interruptions from accidents, are also well described by power laws
\citep{sornette_critical_2006, benguigui_classificacion_2016}.
%
The power law behavior has been observed in a variety of systems,
including for example stock market fluctuations, sizes of computer
files and word frequency in languages \citep{mitzenmacher_brief_2004,
newman_power_2005, simkin_theory_2006}
%
Power laws have also been widely used in biological sciences, e.g., in
analyzing connectivity patterns in metabolic networks
\citep{jeong_large_2000} and in the number of species observed per
unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
% 
More examples can be found in the literature
\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
maccone_evolution_2014, maccone_lognormals_2014,
benguigui_classificacion_2016}

In this work, the exponential distribution of lifespan and waiting
times is justified by considering the hypothesis that the process of
appearance of life in the galaxy is homogeneous and stationary.
%
That is, there is no preferred location within the GHZ for the
spontaneous appearance of life, and the emergence of a \ceti{} is
independent of the existence of previous nodes in the galaxy.
%            
This is equivalent to proposing a Poisson process for the emergence of
\cetis{}, since there is a close relation between the number of events
in time or space and the waiting time or separation, respectively
\citep[e.g., ][]{ross_simulation_2012}.
%
That is, these are two alternative approaches to describing the same
process, a Poisson distribution for the number of events implies an
exponential distribution for their separations, and vice versa.
%
It should be emphasized that the exponential laws used in this work
are assumed as part of the working hypothesis, and instead of
analyzing results from a particular parameter chosen ad--hoc, we
explore the hypothesis space and analyze the impact of the values of
these parameters on the results.
%}}}

\begin{figure} \centering
   \includegraphics[width=0.5\textwidth]{F_number_of_contacts.pdf}
   \usecaption{F_number_of_contacts} \label{F_number_of_contacts}
\end{figure}

\subsection{Complexity of the model}
%{{{

In this Section we discuss the degree of complexity in the model,
considering a compromise between the accuracy of the model and the
number of parameters that are free or with a high uncertainty.
%
Firstly, we emphasize that the odds of a causal contact between two
\cetis{} should not be considered as the odds of a contact between two
intelligent civilizations, and in fact the later could be much lesser
than the former.
%
Indeed, in order to establish a contact between any two entities, a
minimum degree of compatibility must be accomplished without any
previous agreement, making the possibility of a contact with a message
that could be deciphered highly rare \citep[see e.g. ][]{forgan_collimated_2014}.
%
Besides the trade--off between the simplicity and the complexity of
the experiments, further analysis could be performed following this
framework in order to explore possible implications of the results for
more detailed configurations.
%
For example, the communication method (isotropic, collimated,
serendipitous) can affect the observables, making it necessary to
implement a correction factor.
%
Taking this into account, our results regarding the probability of
causal contact should be considered as upper limits for effective
contacts, since they depend on the efficiency of both the emitter and
the receiver to broadcast and scan the sky for intelligent signals,
respectively \citep{grimaldi_signal_2017}.
%
A correction by coverage ratio in the detection and by a targeting
ratio in the emission could be easily implemented in the simulation,
although the effect of reducing the probability of contact is
basically the product of the efficiency ratios and thus such
implementation is not necessary.
%
Therefore, the values of the probabilities could be modified by a
constant correction factor equal to the combined emission/reception
efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019,
forgan_collimated_2014} or for beam--like transmissions
\citep{grimaldi_signal_2017}.
%
Other considerations include the effects of alignments, the use of
stars as sources or amplifiers \citep{Edmondson2003, borra_searching_2012}, 
the nature of the message carrier, the
use of a spatial distribution that actually resembles the spiral shape
and the width of the disc of the Galaxy, different distribution
functions for the mean lifetime of \cetis{}, and different
efficiencies for the sphere of the causal region of each \ceti{}
resembling different searching strategies
\citep{hippke_interstellar_2017}.
%
It is also possible to consider that D$_{max}$ is different for
different \cetis{}.  For example a power law where powerful emissions
are rare and weak emissions are common could be an improvement to the
model.
%
In this work we chose not to implement this for the sake of
simplicity.
%
Finally, the role of the message contents could influence on the
lifespan of a \ceti{} that receives a message, although the
implementation of such behavior would increase the number of free
parameters and would be highly speculative.

%}}}

\subsection{Discrete event process}
%{{{

Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes
the functional forms of the statistical distributions and fixed values
for the hypotheses, a discrete event simulation is performed by
keeping track of a set of variables that change each time a meaningful
event is produced.
%
The main variables that follow the evolution of the simulation are the
positions of stars, which are sampled randomly within the GHZ, the
time of the awakening of each \ceti{} (A event), and the time of the
vanishing of each \ceti{} (D event).
%
The variables that can be deduced from the previous ones include the
number of \cetis{} in casual contact with at least another \ceti{} at a given
time, the number of \cetis{} as a function of time, the number of
\cetis{}
that receive a message at least one time, the number of \cetis{} that
receive a message at least one time and successfully deliver an answer
and the number distribution of waiting time to receive a message.
%
All these quantities are updated each time one of the four events (A,
B, C, D) is produced.

%%% comparar con el stochsatic cellular automata
% ver libro Solving Fermi paradox, sec. 27.3.1

%}}}

\begin{figure*} % 2D color plot
   \centering
   \includegraphics[width=\textwidth]{F_never_contact.pdf}
   \usecaption{F_never_contact}
   \label{F_never_contact}
\end{figure*}

\begin{figure*} % 2D color plot
   \centering
   \includegraphics[width=\textwidth]{F_C_at_A.pdf}
   \usecaption{F_C_at_A}
   \label{F_C_at_A}
\end{figure*}

\subsection{Model for \cetis{}}
%{{{

Since the method is based on events, the first step is to define an
architecture of events, the relationships between them and how events
trigger changes on the state of the system,
so that the system is updated each time an event is produced.
%
We consider a simulated system that represents the galactic habitable
zone which, on a first approach, is a 2-dimensional annular region.
%
The adopted values for the geometry of the GHZ are 20 kly and 60 kly, for the inner
and outer radius, respectively \citep{lineweaver_galactic_2004}.
% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
% mas chico...
%
This simple model does not take into account the variations in stellar
density given by the spiral structure.
%
Although there are several possible approaches, we chose to follow the
evolution of the system according to the following events:

\begin{enumerate}
   \item[(A)] Awakening: A new \ceti{} acquires the ability to emit
      and receive messages and starts to actively broadcast and
      lookout for messages.
   \item[(C)] Contact: A new one-way causal contact is established
      (does not require the awareness of the involved \cetis{}.)
   \item[(B)] Blackout: An existing causal contact is interrupted,
      with or without the occurrence of a doomsday event.
   \item[(D)] Doomsday: An old \ceti{} disappears, or halts its
      communication activities.
\end{enumerate}

We have chosen this nomenclature so that it can be easily remembered
from the initial A, B, C, D.
%
The \blackout is produced when one of the two \cetis{} halts in its
capability of emitting and receiving signals.
%
It can occur at the same time the \doomsday or before.
%
Similarly, a \ccontact can be produced at the time of the \aawakening
or later, and there can be several \contacts for the same node.
%
The time marks for each event (A and D for each \ceti{} and C and B
for each causal contact) are stored as a result of the simulation.
%
Also, the list of active \cetis{} is obtained as a function of time.
%
%Some fixed variables must be set in order to carry out the simulation:
%the size of the Galactic Habitable Zone, the mean lifetime of a
%\ceti{}, $\tau_s$, the mean waiting time for the appearance of another
%\ceti{}, $\tau_a$, and the maximum distance a message can be detected.

The Galactic Habitable Zone is set at fixed values of the inner and outer
radii, radial symmetry is assumed, and the with of the galactic disk
is neglected compared to the radial size.
%
The two radii are measured in light years, with the aim to maintain a
single and comprehensive unit for both space and time coordinates. 
%
The simulation was implemented on a python 3 code, which is publicly
available\footnote{https://github.com/mlares/simu\_contact}

%}}}

\begin{figure}
   \centering
   \includegraphics[width=0.5\textwidth]{F_waiting_for_1C.pdf}
   \usecaption{F_waiting_for_1C}
   \label{F_waiting_for_1C}
\end{figure}

\setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{ccccc}
\hline
   Model subset & D$_{max}$ [lyr] & $\tau_a$ interval [yr] & $\tau_s$ interval 
   [yr]& description  \\
\hline
M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
\hline
\end{tabular}
\caption{Selected models to analyze the behavior of simulation outputs
   and their dependence on simulation parameters.}
\label{T_selected_models}
\end{table*}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Results: exploring the parameter space}\label{S_results}
%{{{

We implemented the simulation of a regular grid of models varying over
the hypothesis space, which covers 15000 different models.
%
For each model, we simulated 50 realizations with different random
seeds, in order to perform a MonteCarlo estimation of the variance.
%
The parameters for the temporal aspects of the simulation (the mean
waiting time for the next awakening, $\left<\tau_a\right>$, and the
mean lifetime, $\left<\tau_s\right>$) cover the ranges 4000-200000 yr
and 10000-500000 yr, respectively, with a regular partition of 50
values for each parameter.
%
For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
20000, 40000 and 80000 lyr.
%
This setup makes a total of 750000 simulations, each covering a time
range of one million years.
%
\ttn{1}
%
In Table \ref{T_simu_hypotheses} we show the three variable
parameters, the ranges of their values and the number of bins that
have been explored in the numerical experiments.
%
We also show the set of parameters that take part in the simulation,
their values and the hypotheses that define the runs of the
simulations.

As a product of the simulations, several quantities can be obtained.
%
Some quantities are directly derived from the discrete events, namely,
the ID of emitting and receiving \cetis{}, the position in the galaxy, and
the times of each of the events that are relevant to keep track of the
number of \cetis{} (time of Awakening and time of Doomsday) and of the
number of contacts (the times of contacts and the times of Blackouts).
%
We can also derive the additional quantities that represent the
properties of the \cetis{}, for example the total time elapsed between the
Awakening and the Doomsday of each node.
%
The time span of a \ceti{} listening another or being listened by another
can also be easily derived from the results of a realization.
%
This way we can also compute the distribution in the galaxy of
\cetis{}
that reach contact and the waiting time until the first contact or the
waiting time until the next contact.
%
Regarding the properties of the population of \cetis{} and its evolution,
the simulation yields the fraction of awaken time a node is listening
at least another node (i.e., in their causal cone), the age of
contacted \cetis{} at first contact, the distribution of time to wait
until next contact, the fraction of \cetis{} where the first contact is
given at the awakening, the distribution of the number of contacts for
each \ceti{}, the distribution of the number of contacts as a function of
the age of the node, the number of contacts as a function of time in the galaxy,
the fractions of nodes that succeed in contact, and the distribution of
distances between contacted nodes.
%
Another useful derived quantity is the duration of two-way
communication channels or the fraction of contacts that admit a
response.
% 
It is also possible to analyze the relations between the distance to
\ceti{} vs. the time of double communication, the distance to \ceti{} vs. the
age of contacted node, the age of a node and the maximum number of
contacted nodes before the Doomsday, or the lifespan of a node vs. the
maximum number of contacts.
%
In addition, all these quantities can be analyzed as a function of the
simulation parameters.
%
We chose eight models that are on fairly opposed regions of the
explored parameter space, and cover short/long lifetimes, short/long
waiting times for the next \ceti{} to appear in the galaxy and short/large
range reach of the signal, including all possible combinations.
%
\ttn{2}
%
The details of each model are summarized in the
Table~\ref{T_selected_models}.

%}}}

\subsection{Membership to the network of connected \cetis{}}\label{SS_members}
%{{{

\ffn{3}
%
In Fig.~\ref{F_number_of_contacts} we show the empirical cumulative
distributions of the number contacts for \cetis{} in six different
samples, including short and long lifetimes
$\left(\,\left<\tau_s\right>=\right.$10000~yr and
$\left<\tau_s\right>$=100000~yr, respectively$\left.\right)$, dense
and sparse spatial distribution
$\left(\,\left<\tau_a\right>=\right.$10000~yr and
$\left<\tau_a\right>$= 100000~yr, respectively$\left.\right)$, and two
different signal reach ranges (D$_{max}$=40000 lyr and
D$_{max}$=80000~lyr).
%
The models M1 and M5 are not shown, since in both cases the total
number of contacts is different from zero for less than the five per
cent of the \cetis{}.
%
The areas between models with the same mean lifetime and mean
awakening time are shaded for visualization purposes.
%
As it can be seen from this Figure, the mean lifetime is more
determinant than the mean awakening rate (dense and sparse,
represented by a different shade) for the number of contacts.
%
As expected, of the samples with the longest $D_{max}$ the model with
a dense awakening in the timeline (low $\left<\tau_a\right>$) and a
long lifetime (high $\left<\tau_s\right>$), is the model that
maximizes the number of contacts, reaching a maximum of order 20
contacts for a single \ceti{}.
%
Similarly, for the models with long signal range
($D_{max}=80000\,lyr$), the one which is dense and with long lifetime
has the maximum number of contacts, reaching nearly 100 contacts for
each \ceti{} in its entire lifetime.
%
This is considerably larger than any model with a shorter lifetime,
which produce a number of contacts of at most the order of ten
contacts per \ceti{}.
%
A couple of comments are worth to mention about this plot, namely, it
has a logarithmic scale on the x-axis, and it is the cumulative, not
differential, empirical distribution, so that the differences in the
number of contacts for different models is quite large.
%
Another interesting feature observed in this Figure, is that for the
models analyzed, there are \cetis{} that have no causal contacts with
any other \ceti{} during their whole lifetime.
%
The fraction of \cetis{} with no contact ranges from nearly 20\% up to
nearly 90\% (models M1 and M5, not shown), except for the dense and
long lifetime models, for which almost all \cetis{} make at least one
contact.
%
Since these results depend on just eight models, we analyze the
properties of the frequency of contacts as a function of both the mean
lifetime and the mean awakening time.
%
\ffn{4}
%
On Fig.~\ref{F_never_contact} we show 2D color maps with the fraction
of \cetis{} in the simulations that never make contact (i.e., never
listen to another \ceti{}), as a function of the mean lifetime
($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
different values of the maximum signal range, $D_{max}=10000$~yr (left
panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
(right panel).
%
A clear pattern emerges in this Figure, showing that the probability
for a \ceti{} of making causal contact with at least another \ceti{}
during their entire lifetime, increases with increasing $D_{max}$,
increasing $\left<\tau_s\right>$ and decreasing $\left<\tau_a\right>$,
following a roughly linear dependence with the three parameters.

While the number of \cetis{} that do not succeed in reaching the
causal contact regions of other node is a useful indicator of
degree of isolation, there is also the chance that a number of
\cetis{} are already in the causal contact region of other nodes,
which shows a somewhat opposite situation.
%
\ffn{5}
%
The fraction of \cetis{} that make the first contact at the awakening
event is shown in the Fig.~\ref{F_C_at_A}, 
as a function of the mean lifetime
($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
different values of the maximum signal range, $D_{max}=10000$~yr (left
panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
(right panel).
%
These probabilities are consistent with the ones on the previous
Figure.
%
Although they are not complementary, it is clear that the dependence
with the three parameters is roughly linear in both cases.
%
%% DISCUTIR ESTOS RESUOLTADOS SEGUN LA LITERATURA

%

%}}}

\subsection{Waiting time for a first contact}\label{SS_waiting}
%{{{

In this subsection we analyze the distribution of the waiting times for
a first contact.
%
Such distribution can be considered to compute the probability for a
random \ceti{} to carry out a SETI project and spend a given time until
the first contact is made.

\ffn{6}
%
In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
waiting times for the first contact, for several models.
%
Upper panel (a) panel shows the histograms for several values of the
mean survival time $\tau_s$, and the mean awakening time $\tau_a$ in
the range 68000-100000 yr.
%
Lower panel (b) shows the histograms for several values of $\tau_a$,
and $\tau_s$ in the range 17000-24000 yr.
%
As it can be seen, the shape of the histograms is almost the same when
$\tau_a$ is modified (panel (b)), but significantly changes when
$\tau_s$ is modified (panel (a)). 
%
For different values of $\tau_a$, the shape of the distribution is
similar, and the main difference is given by the different number of
contacts on the different models.
%
Conversely, for different values of $\tau_a$, the shape of the
distribution is similar, and the main difference given by the
different number of contacts on the different models.

The fraction of the \cetis{} that have made at least one contact as a
function of the elapsed time since the awakening, is computed with
respect to the total number of nodes that make causal contact at least
one time in the time range from the Awakening up to the Doomsday
events.
%
From a frequentist approach, the total counts from the awakening
($t=$0~yr) up to a given time are related to the estimation of the
probability of listening during that time interval and reaching to
causal contact region of at least another node.
%
The complement of this value is the probability of observing in a time
interval with no success, i.e., without ever happening a Contact
event.
%
Clearly, this probability diminishes with time and tends to zero for
large time periods.
%             
The diferential probability of making a causal contact reach its
maximum at the awakening event for all models.
%
That is, for a short period of time like the nearly 50 years SETI
programs have been active, the initial moment is the most promising
for making a (causal) contact, at least for a given technology.
%
This consideration offers a new approach to SETI programs, where the
search for new communication technologies has a fundamental role.
%
Moreover, as it can be seen in the Fig.~\ref{F_waiting_for_1C}, there
is a maximum probability of making the first causal contact at nearly
100000 years.
%
Remarkably, the probability of such a contact does not grows linearly
with the observing time, but is greater during the first 100
millennia.

%}}}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Discussion}\label{S_discussion}
%{{{

We have implemented a suite of simulations following a stochastic
approach, to explore the hypothesis space of a simple model that
accounts for the causal connections between communicating
civilizations on a simplified galaxy.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a galaxy model where an intelligent civilization is
very rare, to a galaxy model populated with plenty of civilizations in
causal contact.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space,
so we explored the outcomes of different models 
that arise as by-products of the simulations.
%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios for
the appearance of intelligent life on the Galaxy.

Although the simulations use a number of speculative assumptions, we
argue that the degree of knowledge about the origin and persistence of
life in the Galaxy does not worth the implementation of more detailed
or sophisticated models.
%
Instead, we take advantage of the simplicity of the model to explore
the hypothesis space, i.e., the three--dimensional parameter space of
the parametric model, in order to gain insight on the consequences of
different scenarios for the search of intelligent life.
%
Our analysis is not centered in obtaining the odds for the Earth
to make contact with another intelligent civilization.
%
Instead, we focus on obtaining a statistical, parameter dependent
description of the possible properties of the communication networks
that comprise sets of nodes with broadcasting and reception
capabilities.
%
This causally connected nodes are sparsely distributed in both space
and time, making it difficult the analytical treatment and justifying
the simulation approach.

Under the hypotheses of our experiments, we conclude that a causal
contact is extremely unlikely unless the galaxy is heavily populated
by intelligent civilizations with large average lifetimes.
%
Roughly, in order to have at least one contact in the entire lifetime
of the \ceti{}, there should appear a mean of at least one \ceti{}
every $\sim$20000 yr, with a mean active period of at least
$\sim$30000 yr.
%
This result is qualitatively similar to the results presented by
several authors, which state that a contact between the Earth and
another intelligent civilization in the Galaxy is quite unlikely,
provided the maximum distance of the signal and the lifetime of the
emitter are not large enough.
%
This analysis supports the idea that, in order to increase the
possibilities of a contact, more active strategies of the emitter
would be required.
%
Some proposals in this direction include intertellar exploration,
colonization and settlement \citep{brin_great_1983, Dosovic2019,
galera_invasion_2019}, although it would require large temporal
scales.
%
\citet{Dosovic2019} use probabilistic cellular automaton simulations
to explore the parameter space of a model with colonization and
catastrophic events.
%
According to the timescales involved, the results could explain the
Fermi paradox.
%
Although our work does not take into account the colonization
hypothesis, it does consider catastrophic events implicitly in the
mean value and distribution of the lifetime, $\left<\tau_s\right>$.
%
Other strategies could also increase the probability of contacts, for
example panspermia \citep[e.g.,][]{starling_virulence_2013} or
self--replicating probes \citep[e.g.,][]{barlow_galactic_2013},
although they would be too slow to make a significant impact on the
communication network among intelligent civilizations.
%
Our results are also consistent with those presented by
\citet{grimaldi_signal_2017}, who estimate an upper bound for the mean
number of extraterrestrial civilizations that could contact Earth
using Monte Carlo simulations, from a statistical model where the with
of the Galactic disk is not negligible.
%
On the other hand, taking into account the probability of contact with
radio signals not sent intentionally \citep{horvat_calculating_2007},
our analysis would imply a highly populated solar neighborhood.
%
Remarkably, most of the studies using statistical models or
simulations are based on the Drake equation
\citep{cirkovic_temporal_2004, smith_broadcasting_2009,
bloetscher_using_2019}.
%
Our approach does not relay on the Drake equation, and thus is not
motivated by the detailed physical processes that give rise to
intelligent life.
%
However, we argue that it is a valid empirical formulation to discuss
the probabilities of contact, and mainly time scales involed in the
problem.
%
Ours is an alternative to the method proposed by
\citet{balbi_impact_2018}, who perform an analysis based on the Earth
and use a different model, although our results agree in general.

%%%%%%%%%%%%%%%

There is a balance between the density of \cetis{} and the mean
lifetime since, as expected, a lower density can be compensated by a
longer active time period.
%
However, a large number of nodes does not easily compensate their
short lives to reach the same probability of causal contact than in
the case of a less populated galaxy but with very ancient
civilizations.
%
In all cases, for a short period of time (for instance, the time SETI
programs have been active on Earth), the maximum probability of making
a contact occurs at the moment of the awakening.
%
This suggests the possibility that an alternative SETI strategy could
be the search for alternative message carriers, for the case in which
the search has not been performed on the adequate channels.
%
Then, if a contact is produced for the first time, the origin of the
signal is more likely to be very old.
%
Also, the chances of entering the causal connected zone of a \ceti{}
does not grow linearly, but favours the first period of 100000 years.

In the approach of this work, we used computer simulations to address
the problem of the probabilities of causal contacts between location
in the galaxy the possibility of sending and receiving messages.
%
Instead of making a number of assumptions, we have explored the
hypothesis space, reducing the problem to only three parameters and a
few simple hypothesis to perform a complete model for the population and
communication network in the galaxy.
%
This allows to consider the Fermi paradox from a new perspective, and
to propose an alternative treatment for the number of intelligent
emitter/receivers with emphasis on the temporal dimension which limits
the probabilities of contacts due to the short time interval for
rise and fall of civilizations compared to the age and extension of
our Galaxy.

%}}}

\ack[Acknowledgement]{
%
This work was partially supported by the Consejo Nacional de
Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
Argentina.
%
This research has made use of NASA's Astrophysics Data System.
%
Plots and simulations were made with software developed by the authors
using R and python languajes. Plots were postprocessed with inkscape.
%
}

\ack[Disclosure statement]{
%
No competing financial interests exist.
%
}

%\bibliographystyle{mn2e}
\setlength{\bibsep}{0.0pt}

\begin{thebibliography}{90}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{Adamic(2000)}]{adamic_zipf_2000}
Adamic L., 2000, {Zipf, Power-laws, and Pareto - a ranking tutorial}. web

\bibitem[{Anchordoqui \& Weber(2019)}]{anchordoqui_upper_2019}
Anchordoqui L.~A., Weber S.~M., 2019, arXiv e-prints

\bibitem[{Annis(1999)}]{annis_astrophysical_1999}
Annis J., 1999, JBIS - Journal of the British Interplanetary Society, 52, 19

\bibitem[{Armstrong \& Sandberg(2013)}]{Armstrong2013}
Armstrong S., Sandberg A., 2013, Acta Astronautica, 89, 1

\bibitem[{Balb(2018)}]{balb_spatiotemporal_2018}
Balb A., 2018, Mem. della Soc. Astron. Ital., 89, 425

\bibitem[{Balbi(2018)}]{balbi_impact_2018}
Balbi A., 2018, Astrobiology, 18, 54

\bibitem[{Barab{\'{a}}si(2009)}]{barabasi_scale_2009}
Barab{\'{a}}si A.~L., 2009, Science (80-. )., 325, 412

\bibitem[{Barlow(2013{\natexlab{a}})}]{barlow_galactic_2012}
Barlow M.~T., 2013{\natexlab{a}}, Int. J. Astrobiol., 12, 63

\bibitem[{Barlow(2013{\natexlab{b}})}]{barlow_galactic_2013}
Barlow M.~T., 2013{\natexlab{b}}, Int. J. Astrobiol., 12, 63

\bibitem[{Benguigui \& Marinov(2016)}]{benguigui_classificacion_2016}
Benguigui L., Marinov M., 2016, {A classification of the natural and social
  distributions Part 2: the explanations}. Tech. rep.

\bibitem[{Bloetscher(2019)}]{bloetscher_using_2019}
Bloetscher F., 2019, Acta Astronaut., 155, 118

\bibitem[{Borra(2012)}]{borra_searching_2012}
Borra E.~F., 2012, Astronomical Journal, 144

\bibitem[{Brin(1983)}]{brin_great_1983}
Brin G., 1983, Quarterly Journal of the Royal Astronomical Society, 24, 283

\bibitem[{Burchell(2006)}]{burchell_whither_2006}
Burchell M.~J., 2006, Int. J. Astrobiol., 5, 243

\bibitem[{Carroll-Nellenback {et~al}\mbox{.}(2019)Carroll-Nellenback, Frank,
  Wright, \& Scharf}]{carroll_nellemback_fermi_2019}
Carroll-Nellenback J., Frank A., Wright J., Scharf C., 2019, The Astronomical
  Journal, 158, 117

\bibitem[{Chung(2003)}]{chung_simulation_2003}
Chung C.~A., 2003, {Simulation Modeling Handbook: A Practical Approach
  (Industrial and Manufacturing Engineering Series)}. CRC press

\bibitem[{{\'{C}}irkovi{\'{c}}(2004)}]{cirkovic_temporal_2004}
{\'{C}}irkovi{\'{c}} M.~M., 2004, Astrobiology, 4, 225

\bibitem[{{Conway Morris}(2018)}]{conway_three_2018}
{Conway Morris} S., 2018, Int. J. Astrobiol., 17, 287

\bibitem[{Dayal {et~al}\mbox{.}(2016)Dayal, Ward, \&
  Cockell}]{dayal_habitability_2016}
Dayal P., Ward M., Cockell C., 2016, arXiv e-prints

\bibitem[{Do{\v{s}}ovi{\'{c}} {et~al}\mbox{.}(2019)Do{\v{s}}ovi{\'{c}},
  Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}}]{Dosovic2019}
Do{\v{s}}ovi{\'{c}} V., Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2019,
  Astronomy and Astrophysics, 625

\bibitem[{Drake(1962)}]{drake_intelligent_1962}
Drake F., 1962

\bibitem[{Edmondson \& Stevens(2003)}]{Edmondson2003}
Edmondson W.~H., Stevens I.~R., 2003, International Journal of Astrobiology, 2,
  231

\bibitem[{Enriquez {et~al}\mbox{.}(2017)Enriquez, Siemion, Foster, Gajjar,
  Hellbourg, Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
  Werthimer}]{enriquez_breakthrough_2017}
Enriquez J.~E. {et~al.}, 2017, Astrophys. J., 849, 104

\bibitem[{Fogg(1987)}]{fogg_temporal_1987}
Fogg M.~J., 1987, Icarus, 69, 370

\bibitem[{Forgan {et~al}\mbox{.}(2016)Forgan, Dayal, Cockell, \&
  Libeskind}]{forgan_evaluating_2015}
Forgan D., Dayal P., Cockell C., Libeskind N., 2016, Int. J. Astrobiol., 16, 60

\bibitem[{Forgan(2009)}]{forgan_numerical_2009}
Forgan D.~H., 2009, Int. J. Astrobiol., 8, 121

\bibitem[{Forgan(2011)}]{forgan_spatiotemporal_2011}
Forgan D.~H., 2011, Int. J. Astrobiol., 10, 341

\bibitem[{Forgan(2013)}]{forgan_possibility_2013}
Forgan D.~H., 2013, JBIS - J. Br. Interplanet. Soc., 66, 144

\bibitem[{Forgan(2014)}]{forgan_collimated_2014}
Forgan D.~H., 2014, JBIS - J. Br. Interplanet. Soc., 67, 232

\bibitem[{Forgan(2017{\natexlab{a}})}]{forgan_galactic_2016}
Forgan D.~H., 2017{\natexlab{a}}, Int. J. Astrobiol., 16, 349

\bibitem[{Forgan(2017{\natexlab{b}})}]{forgan_galactic_2017}
Forgan D.~H., 2017{\natexlab{b}}, Int. J. Astrobiol., 16, 349

\bibitem[{Forgan(2019)}]{forgan_exoplanet_2017}
Forgan D.~H., 2019, Int. J. Astrobiol., 18, 189

\bibitem[{Forgan \& Rice(2010)}]{forgan_numerical_2010}
Forgan D.~H., Rice K., 2010, Int. J. Astrobiol., 9, 73

\bibitem[{Frank(2009)}]{frank_common_2009}
Frank S.~A., 2009, {The common patterns of nature}

\bibitem[{Funes {et~al}\mbox{.}(2019)Funes, Florio, Lares, \&
  Asla}]{funes_searching_2019}
Funes J.~G., Florio L., Lares M., Asla M., 2019, Theol. Sci., 0, 1

\bibitem[{Galera {et~al}\mbox{.}(2019)Galera, Galanti, \&
  Kinouchi}]{galera_invasion_2019}
Galera E., Galanti G.~R., Kinouchi O., 2019, International Journal of
  Astrobiology, 18, 316

\bibitem[{Glade {et~al}\mbox{.}(2012)Glade, Ballet, \&
  Bastien}]{glade_stochastic_2011}
Glade N., Ballet P., Bastien O., 2012, Int. J. Astrobiol., 11, 103

\bibitem[{Gobat \& Hong(2016)}]{gobat_evolution_2016}
Gobat R., Hong S.~E., 2016, Astron. Astrophys., 592

\bibitem[{Gonzalez(2005)}]{gonzalez_habitable_2005}
Gonzalez G., 2005, Orig. Life Evol. Biosph., 35, 555

\bibitem[{Gonzalez {et~al}\mbox{.}(2001)Gonzalez, Brownlee, \&
  Ward}]{gonzalez_galactic_2001}
Gonzalez G., Brownlee D., Ward P., 2001, Icarus, 152, 185

\bibitem[{Grimaldi(2017)}]{grimaldi_signal_2017}
Grimaldi C., 2017, Sci. Rep., 7, 46273

\bibitem[{Grimaldi {et~al}\mbox{.}(2018)Grimaldi, Marcy, Tellis, \&
  Drake}]{Grimaldi2018}
Grimaldi C., Marcy G.~W., Tellis N.~K., Drake F., 2018, Publications of the
  Astronomical Society of the Pacific, 130

\bibitem[{Haqq-Misra(2019)}]{haqq-misra_evolution_2019}
Haqq-Misra J., 2019, arXiv e-prints

\bibitem[{Haqq-Misra \& Kopparapu(2018)}]{haqq-misra_drake_2017}
Haqq-Misra J., Kopparapu R.~K., 2018, Habitability of the Universe Before
  Earth, 307

\bibitem[{Harp {et~al}\mbox{.}(2018)Harp, Ackermann, Astorga, Arbunich,
  Barrios, Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak,
  \& Tarter}]{harp_application_2018}
Harp G.~R. {et~al.}, 2018, Astrophys. J., 869, 66

\bibitem[{Hart(1975)}]{hart_explanation_1975}
Hart M.~H., 1975, Quarterly Journal of the Royal Astronomical Society, 16, 128

\bibitem[{Hinkel {et~al}\mbox{.}(2019)Hinkel, Hartnett, Lisse, \&
  Young}]{hinkel_interdisciplinary_2019}
Hinkel N., Hartnett H., Lisse C., Young P., 2019, Bull. Am. Astron. Soc., 51,
  497

\bibitem[{Hippke(2017)}]{hippke_interstellar_2017}
Hippke M., 2017, arXiv e-prints

\bibitem[{Horvat(2006)}]{horvat_calculating_2007}
Horvat M., 2006, Int. J. Astrobiol., 5, 143

\bibitem[{Horvat {et~al}\mbox{.}(2011)Horvat, Naki{\'{c}}, \&
  Oto{\v{c}}an}]{horvat_impact_2011}
Horvat M., Naki{\'{c}} A., Oto{\v{c}}an I., 2011, Int. J. Astrobiol., 11, 51

\bibitem[{Jeong {et~al}\mbox{.}(2000)Jeong, Tombor, Albert, Oltvai, \&
  Barabasi}]{jeong_large_2000}
Jeong H., Tombor B., Albert R., Oltvai Z.~N., Barabasi A.~L., 2000, Nature,
  407, 651

\bibitem[{Lampton(2013)}]{lampton_information_2013}
Lampton M., 2013, Int. J. Astrobiol., 12, 312

\bibitem[{Lineweaver {et~al}\mbox{.}(2004)Lineweaver, Fenner, \&
  Gibson}]{lineweaver_galactic_2004}
Lineweaver C.~H., Fenner Y., Gibson B.~K., 2004, Science (80-. )., 303, 59

\bibitem[{Lingam \& Loeb(2018)}]{lingam_relative_2019}
Lingam M., Loeb A., 2018, Astrobiology, 19, 28

\bibitem[{Loeb {et~al}\mbox{.}(2016)Loeb, Batista, \&
  Sloan}]{loeb_relative_2016}
Loeb A., Batista R.~A., Sloan D., 2016, J. Cosmol. Astropart. Phys., 2016, 040

\bibitem[{Loeb \& Zaldarriaga(2007)}]{loeb_eavesdropping_2006}
Loeb A., Zaldarriaga M., 2007, J. Cosmol. Astropart. Phys., 1

\bibitem[{Maccone(2010)}]{maccone_KLT_2010}
Maccone C., 2010, Acta Astronaut., 67, 1427

\bibitem[{Maccone(2011)}]{maccone_mathematical_2011}
Maccone C., 2011, Orig. Life Evol. Biosph., 41, 609

\bibitem[{MacCone(2011)}]{maccone_SETI_2011}
MacCone C., 2011, Acta Astronaut., 68, 63

\bibitem[{MacCone(2013)}]{maccone_SETI_2013}
MacCone C., 2013, Int. J. Astrobiol., 12, 218

\bibitem[{Maccone(2014{\natexlab{a}})}]{maccone_evolution_2014}
Maccone C., 2014{\natexlab{a}}, Int. J. Astrobiol., 13, 290

\bibitem[{Maccone(2014{\natexlab{b}})}]{maccone_lognormals_2014}
Maccone C., 2014{\natexlab{b}}, Acta Astronaut., 105, 538

\bibitem[{Maccone(2015)}]{maccone_statistical_2015}
Maccone C., 2015, Acta Astronaut., 115, 277

\bibitem[{Mart{\'{i}}n \& Goldenfeld(2006)}]{martin_origin_2006}
Mart{\'{i}}n H.~G., Goldenfeld N., 2006, Proc. Natl. Acad. Sci. U. S. A., 103,
  10310

\bibitem[{Mitzenmacher(2004)}]{mitzenmacher_brief_2004}
Mitzenmacher M., 2004, Internet Math., 1, 226

\bibitem[{Morrison \& Gowanlock(2015)}]{morrison_extending_2015}
Morrison I.~S., Gowanlock M.~G., 2015, Astrobiology, 15, 683

\bibitem[{Murante {et~al}\mbox{.}(2015)Murante, Monaco, Borgani, Tornatore,
  Dolag, \& Goz}]{murante_simulating_2015}
Murante G., Monaco P., Borgani S., Tornatore L., Dolag K., Goz D., 2015, Mon.
  Not. R. Astron. Soc., 447, 178

\bibitem[{Newman(2005)}]{newman_power_2005}
Newman M., 2005, Contemp. Phys., 46, 323

\bibitem[{Newman \& Sagan(1981)}]{newman_galactic_1981}
Newman W.~I., Sagan C., 1981, Icarus, 46, 293

\bibitem[{Peters(2018)}]{peters_outer_2018}
Peters T., 2018, Int. J. Astrobiol., 17, 282

\bibitem[{Prantzos(2013)}]{prantzos_joint_2013}
Prantzos N., 2013, Int. J. Astrobiol., 12, 246

\bibitem[{Ptolemaeus(2014)}]{ptolemaeus_system_2014}
Ptolemaeus C., 2014, {System Design, Modeling, and Simulation. Using Ptolemy
  II}. Ptolemy.org, Berkeley, p. 690

\bibitem[{Rahvar(2017)}]{rahvar_cosmic_2016}
Rahvar S., 2017, Mon. Not. R. Astron. Soc., 470, 3095

\bibitem[{Ramirez {et~al}\mbox{.}(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz,
  V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez}]{ramirez_new_2017}
Ramirez R., G{\'{o}}mez-Mu{\~{n}}oz M.~A., V{\'{a}}zquez R., N{\'{u}}{\~{n}}ez
  P.~G., 2017, International Journal of Astrobiology, 1

\bibitem[{Ross(2012)}]{ross_simulation_2012}
Ross S.~M., 2012, {Simulation}. Elsevier Science Publishing Co Inc

\bibitem[{Simkin \& Roychowdhury(2006)}]{simkin_theory_2006}
Simkin M.~V., Roychowdhury V.~P., 2006, J. Math. Sociol., 30, 33

\bibitem[{Smith(2009)}]{smith_broadcasting_2009}
Smith R.~D., 2009, Int. J. Astrobiol., 8, 101

\bibitem[{Solomonides {et~al}\mbox{.}(2016)Solomonides, Kaltenegger, \&
  Terzian}]{solomonides_probabilistic_2016}
Solomonides E., Kaltenegger L., Terzian Y., 2016, 228th AAS, San Diego, 228, 1

\bibitem[{Sornette(2006)}]{sornette_critical_2006}
Sornette D., 2006, {Critical phenomena in natural sciences: Chaos, fractals,
  selforganization and disorder: Concepts and tools}. Springer

\bibitem[{Sotos(2019)}]{Sotos_biotechnology_2019}
Sotos J.~G., 2019, {Biotechnology and the lifetime of technical civilizations}

\bibitem[{Starling \& Forgan(2014)}]{starling_virulence_2013}
Starling J., Forgan D.~H., 2014, Int. J. Astrobiol., 13, 45

\bibitem[{Tarter(2001)}]{tarter_search_2001}
Tarter J., 2001, Annu. Rev. Astron. Astrophys., 39, 511

\bibitem[{Tarter {et~al}\mbox{.}(2009)Tarter, Backus, Blair, Cordes, Harp,
  Henry, Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
  Werthimer}]{tarter_advancing_2009}
Tarter J. {et~al.}, 2009, Astro2010 Astron. Astrophys. Decad. Surv. Sci. White
  Pap. no. 294, 2010

\bibitem[{Vakoch \& Dowd(2015)}]{vakoch_drake_2015}
Vakoch D.~A., Dowd M.~F., 2015, {The drake equation: Estimating the prevalence
  of extraterrestrial life through the ages}, Vakoch D.~A., Dowd M.~F., eds.
  Cambridge University Press, Cambridge, pp. 1--319

\bibitem[{VanHouten(2018)}]{vanhouten_isthere_2017}
VanHouten M.~A., 2018, Lancet Gastroenterol. Hepatol., 3, 382

\bibitem[{Vukoti{\'{c}} \&
  {\'{C}}irkovi{\'{c}}(2012)}]{vukotic_astrobiological_2012}
Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2012, Orig. Life Evol. Biosph.,
  42, 347

\bibitem[{Vukoti{\'{c}} {et~al}\mbox{.}(2016)Vukoti{\'{c}}, Steinhauser,
  Martinez-Aviles, {\'{C}}irkovi{\'{c}}, Micic, \&
  Schindler}]{vukotic_grandeur_2016}
Vukoti{\'{c}} B., Steinhauser D., Martinez-Aviles G., {\'{C}}irkovi{\'{c}}
  M.~M., Micic M., Schindler S., 2016, Mon. Not. R. Astron. Soc., 459, 3512

\bibitem[{Walters {et~al}\mbox{.}(1980)Walters, Hoover, \&
  Kotra}]{walters_interstellar_1980}
Walters C., Hoover R.~A., Kotra R., 1980, Icarus, 41, 193

\bibitem[{Wright {et~al}\mbox{.}(2015)Wright, Cartier, Zhao, Jontof-Hutter, \&
  Ford}]{wright_theGsearch_2015}
Wright J.~T., Cartier K. M.~S., Zhao M., Jontof-Hutter D., Ford E.~B., 2015,
  Astrophys. J., 816, 17

\bibitem[{Wright {et~al}\mbox{.}(2018)Wright, Kanodia, \&
  Lubar}]{wright_how_2018}
Wright J.~T., Kanodia S., Lubar E., 2018, The Astronomical Journal, 156, 260

\end{thebibliography}

\end{document}